### 2014-10-28 mswp-proposal.tex
Initial commit

\documentclass[a4paper,11pt]{article}

\usepackage{graphicx}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% \usepackage{mathptmx} % Times New Roman

\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\footskip}{15mm}

\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}

% --------------------------------------------------------------

\providecommand*{\ShortTitle}{$<$Short title$>$}
\providecommand*{\FullTitle}{$<$Full title$>$}

% --------------------------------------------------------------

\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
{\textbf{\sffamily\Large \FullTitle}}
\vspace{1cm}}

\author{
{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
Group: $<$GID$>$\bigskip \\
$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
\vspace{4cm}
}

\begin{document}

\begin{titlepage}
\maketitle

\end{titlepage}

% --------------------------------------------------------------

\thispagestyle{empty}
\tableofcontents
\pagebreak

\setcounter{page}{1}

% --------------------------------------------------------------

\note{
\textbf{Formal constraints}
\begin{itemize}
\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
\item    Fontsize: 11pt
\item     Single line spacing
\item     Margins: 2.5cm side and top/bottom
\item     \fbox{Language: ENGLISH}
\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
\end{itemize}
\textbf{Available templates}
\begin{itemize}
\item     Proposal (mswp-proposal.tex)
\item     Costs (costs.xls, costs.ods)
\end{itemize}
\textbf{Supplemental material}
\begin{itemize}
\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
\item     Ethical issues form (ethical-issues.rtf)
\end{itemize}
}
\pagebreak

% --------------------------------------------------------------
\section{Synopsis}
\label{sect:synopsis}

\note{
\begin{itemize}
	\item {\em Length: 1-2 pages}
	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach. 
	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
	\item Short description of project idea.
	\item Short description of problem / situation.
	\item What makes it an interesting endeavor?
	\item Why should somebody care?
	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
	\item Who are the beneficiaries of the results?
	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
	\item Constraints
	\begin{itemize}
		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
		\item     The project should be interesting and meaningful.
		\item     {\em Your creativity is called for!!}
	\end{itemize}
\end{itemize}
}

% --------------------------------------------------------------
\section{Introduction and problem description}
\label{sect:intro}

\note{
\begin{itemize}
	\item {\em Length: 2-3 pages}
	\item {\bf Why?}
	\item Introduction
	\item Context
	\item What is the current situation?
	\item What is the open/unresolved problem or opportunity?
	\item Why is it a problem?
	\item What is unknown?
	\item What could be improved?
	\item Explanation of fundamental terms and basic definitions.
\end{itemize}
}

% --------------------------------------------------------------
\section{Project goals and deliverables}
\label{sect:goals}

\note{
\begin{itemize}
	\item {\em Length: 1-2 pages}
	\item What is the goal of the project?
	\item Research questions
	\begin{itemize}
		\item 	    What are the hypotheses that are to be investigated?
		\item 	    Main hypothesis \& sub hypotheses
	\end{itemize}
	\item Which results should be achieved with the project?
	\begin{itemize}
	    	\item 	   What will be known afterwards that is not known now?
		\item	    What will be created that does not exist now?
	\end{itemize}
	\item Non-goals (What will not be part of the project? What will not be done?)
\end{itemize}
}

% --------------------------------------------------------------
\section{Scientific relevance and innovative aspects}
\label{sect:relevance}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Why is the project scientifically interesting?
\item Did others point out that this is an open question?
\item What are the innovative aspects that make it interesting?
\item How could the project break new ground scientifically?
\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
\end{itemize}
}

% --------------------------------------------------------------
\section{State of the art / current knowledge}
\label{sect:star}

\note{
\begin{itemize}
\item {\em Length: 2-5 pages}
\item What results and approaches have already been presented in this or related areas?
\item Relation to the international scientific work in the field (international status of the research)
\item Description and critical discussion of related scientific work
\end{itemize}
}

% --------------------------------------------------------------
\section{Method}
\label{sect:method}

\note{
\begin{itemize}
\item {\em Length: 2-5 pages}
\item {\bf How?}
\item How should the expected results be achieved?
\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
\item Description of the methods.
\item Justifications for chosen methods.
\end{itemize}
}

% --------------------------------------------------------------
\section{Detailed description of the workpackages}
\label{sect:workplan}

\note{
\begin{itemize}
\item {\em Length: 2-4 pages}
\item Structuring the project into self-contained parts.
\item Additional verbal descriptions.
\item Work packages
    \begin{itemize}
    \item title
    \item goal(s)
    \item description
    \item expected results
    \item responsible person(s)
    \item dependencies
    \end{itemize}
\end{itemize}
}

% --------------------------------------------------------------
\section{Time plan (Gantt chart)}
\label{sect:timeplan}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Realistic estimation of schedule based on workpackages.
\item Including milestones (not only when but also what is to be achieved for each milestone).
\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
\end{itemize}
}

% --------------------------------------------------------------
\section{Human resources / team}
\label{sect:team}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
\item How many people?
\item To what extent are individual members needed?
\item What knowledge, skills, and experiences are needed for each member?
\item Demonstrate that the members will be able to carry out the project successfully.
\item Work structure
	\begin{itemize}
	\item     Who will lead the project?
	\item     How do they work together?
	\item     Management and coordination
		\begin{itemize}
		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
		\item 	        How often will meetings take place? (Who will participate?)
		\item 	        How will the work be documented?
		\item 	        How will information be stored and shared?
		\end{itemize}
	\end{itemize}
\item Cooperations
	\begin{itemize}
	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
	\item     What is their role?
	 \item    Why are they needed?
	\end{itemize}
\end{itemize}
}

% --------------------------------------------------------------
\section{Costs}
\label{sect:costs}

\note{
\begin{itemize}
\item {\em Length: 2-3 pages}
\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
\item An Excel template is provided as supplementary material to support budgeting.
\item Personnel costs
	\begin{itemize}
	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.). The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB). In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
	 \item    The justification of the requested personnel should contain:
		\begin{itemize}
		\item 	        description of type of work;
		\item 		        extent of involvement (part-time contracts are permitted).
		\end{itemize}
	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
	\end{itemize}
\item Equipment costs
	\begin{itemize}
	\item     Indicate reasons for equipment costs. The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
	\end{itemize}
\item Material costs
	\begin{itemize}
	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT. The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan. Experience with previous projects should be taken into account.
	\end{itemize}
\item Travel costs
	\begin{itemize}
	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc. Applicants are to provide a detailed travel (cost) plan broken down by project participant. For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV). The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}). For longer stays an appropriate and comprehensible cost plan should be prepared.
	\end{itemize}
\item Other costs
	\begin{itemize}
	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
		\begin{itemize}
		\item         reimbursement of costs towards or for the use of research facilities, e.g. of large-scale research facilities (project-specific 'equipment time'). Applicants should obtain and submit multiple offers;
		\item         costs for project-specific work carried out outside the applicant's research institution (e.g. for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.). Applicants should obtain and submit multiple offers;
		\item         honoraria for test persons;
		\end{itemize}
	\end{itemize}
\end{itemize}
}

% --------------------------------------------------------------
\section{Expected implications and risks}
\label{sect:implication-risk}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Importance of the expected results for the discipline
	\begin{itemize}
	\item     To what extent does the proposed research address important challenges?
	\end{itemize}
\item Importance of the expected results for other areas
\item What are possible risks of the project and how can they be alleviated?
	\begin{itemize}
	\item     What factors could lead to a failure of the project?
	\item     Which factors or persons could support the project and increase the chance for success?
	\item     What if important team members leave the project?
	\end{itemize}
\end{itemize}
}

% --------------------------------------------------------------
\section{Ethical considerations \& security issues}
\label{sect:ethics-security}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
\item Are there any security-sensitive issues that apply to your proposal?
\end{itemize}
}

% --------------------------------------------------------------
% APPENDIX
\begin{appendix}

\pagebreak

% --------------------------------------------------------------
% References
\phantomsection
\addcontentsline{toc}{section}{References}

\bibliographystyle{apalike}
\bibliography{mswp2013-proposal}

\pagebreak

% --------------------------------------------------------------
% Abbreviations
\section*{Abbreviations}
 \addcontentsline{toc}{section}{Abbreviations}

 \begin{description}
  \item[MSWP] Management von Software Projekten
  \item[WP] Work Package
 \end{description}

\end{appendix}

\end{document}

### 2014-10-31 diary/10_Diary_1.tex
adds diary structure & raw report about the first two sessions

--- 2014-10-28 mswp-proposal.tex
@@
 
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
+\providecommand*{\ShortTitle}{WaveMeIn}
+\providecommand*{\FullTitle}{Projektprotokoll}
+\providecommand*{\includeWiki}[1]{\wikiEnvironments \include{#1}\nowikiEnvironments}
 
@@
 {\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
+Group: 10\bigskip \\
+Belk Stefan \\ {\small 0750926, 937, \href{mailto:belk.stefan@gmail.com}{belk.stefan@gmail.com}}\\
+Petz Thomas \\ {\small 0601280, 937, \href{mailto:e0601280@student.tuwien.ac.at}{e0601280@student.tuwien.ac.at}}\\
+Causevic Alma \\ {\small 0847805, 534, \href{mailto:alma.causevic@hotmail.com}{alma.causevic@hotmail.com}}\\ 
+Causevic Amra  \\ {\small 0649241, 534, \href{mailto:amra.causevic@hotmail.com}{amra.causevic@hotmail.com}}\\ 
+Seebacher David \\ {\small 0327243, 534, \href{mailto:david.seebacher@student.tuwien.ac.at}{david.seebacher@student.tuwien.ac.at}}\\
 \vspace{4cm}
@@
 
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
+%\input{diary/28.10.md}
+%\input{diary/31.10.md}
+\input{diary_bundle.tex}
 \end{document}

### 2014-10-31 diary/diary_bundle.tex
adds diary structure & raw report about the first two sessions

\section{Erstes Treffen - Di 28 Okt
2014}\label{erstes-treffen---di-28-okt-2014}

\begin{itemize}
\item
  \textbf{Anwesende:} alle
\item
  \textbf{Abwesend:} niemand
\item
  \textbf{Ort:} Freihaus, Arbeitsraum gelb
\item
  \textbf{Zeit:} 17:45-20:10
\item
  \textbf{Tagesordnung:} kennenlernen, Rahmen der LVA besprechen,
  Unterlagen sichten und verteilen
\item
  Unterlagen und Vorlagen bereitstellen

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Dropboxordner wird erstellt
  \item
    git-Repository wird erstellt
  \item
    alle Vorlagen werden bereitgestellt
  \end{itemize}
\item
  Brainstorming

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    eRezept (ähnlich eMedikation)
  \item
    Goatsimulator (Gedanken verstehen)
  \item
    App für freie Parkplätze
  \item
    Passwörter über Gehirnwellen eingeben
  \item
    Stalking App (??)
  \item
    Kleidung nach Wetter und Laune (intelligenter Kleiderschrank)
  \item
    Wahlapp (wählt von allein?)
  \item
    Sichere Email Alternative
  \item
    Dioptrin messen / Optiker App
  \item
    Intelligente Taschentücher (Gesundheitszustand)
  \item
    Schnitzelscanner (Essen scannen, Kalorien, Zutaten)
  \item
    Semantic porn/shoes (oder, wer ist gut in Mathe?)
  \item
    OnlineStudentManager
  \end{itemize}
\item
  4 finale Ideen

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    OnlineStudentManager
  \item
    eRezept
  \item
    BrainCode
  \item
    WhoKnowsWhat
  \end{itemize}
\item
  Eine Mail mit den Vorschlägen wird dem Tutor geschickt
\item
  Planung des nächsten Treffens: Freitag 31.10, 12 Uhr im gleichen Raum.
  Thema. Feedback des Tutors, Ausarbeitung der Grundlagen des Projektes
  und Vorbereitung der Folien
\end{itemize}

\section{Grundlagen des Projektes klären - Fr 31 Okt
2014}\label{grundlagen-des-projektes-kluxe4ren---fr-31-okt-2014}

\begin{itemize}
\item
  \textbf{Anwesende:} alle
\item
  \textbf{Abwesend:} niemand
\item
  \textbf{Ort:} Freihaus, Arbeitsraum gelb
\item
  \textbf{Zeit:} 12:10-
\item
  \textbf{Tagesordnung:} Feedback des Tutors besprechen, Ausarbeitung
  der Grundlagen des Projektes und Vorbereitung der Folien
\item
  Was muss für 6.11 vorbereitet werden?

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Präsentation von 7 Minuten
  \item
    Präsentationsfolien in GoogleDocs vorbereitet
  \item
    Projektidee formulieren
  \item
    lustiges Video über Gehirnwellen-gesteuerte Geräte am Ende zeigen
    wenn Zeit bleibt
  \end{itemize}
\item
  Diskussion wie das Erkennen der Gedanken funktionieren könnte

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    http://www.dailymail.co.uk/sciencetech/article-2095214/As-scientists-discover-translate-brainwaves-words--Could-machine-read-innermost-thoughts.html
  \item
    Ist eine persönliche Analyse relevant? Muss man das Gerät
    trainieren?

    \begin{itemize}
    \itemsep1pt\parskip0pt\parsep0pt
    \item
      mehrere Personen sollen mit verschiedenen Profilen unterstützt
      werden
    \item
      ist es möglich für mehrere Benutzer bei
    \end{itemize}
  \end{itemize}
\item
  Diskussion wie das Entsperren/Passworteingabe funktionieren könnte

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Es ist sinnvoll ein Masterpasswort
  \end{itemize}
\item
  Diskussion über Stromversorgung und Geräteform

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    Bluetoothheadset ist wegen Akku so groß
  \item
    Hörgeräte sind mittlerweile relativ kein
  \item
    Strominduktion durch Radiowellen (genug Energie?, legal?)
  \item
    Stromversorgung durch Bewegung (Pendel..)
  \item
    logcost EEG-Sensor
  \item
    Sind maßgeschneiderte Geräte sinnvoll?
  \item
    eventuell in Bluetoothgeräte einbauen
  \item
    an Brillen befestigen (wie Googleglasses)
  \item
    Muss an der Stirn abgenommen werden?
  \end{itemize}
\item
  Diskussion über generelle Sinnhaftigkeit und Alternativen

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    es ist praktisch
  \item
    es kann von Menschen verwendet werden die alternativen wie
    Fingerabdrucksensor oder Retinascans nicht verwenden können
  \item
    Fingerabdrücke und Retinascanns (Biometrie generell) sind nach
    Kompromittierung permanent unsicher
  \item
    Verschmutzung des Scanners nicht tragisch
  \end{itemize}
\item
  Diskusion über die Sicherheit der Verbindung

  \begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item
    ist es Hackbar?
  \item
    wie sicher ist Bluetoothübertragung? Muss recherchiert werden
  \item
    Sind Gehirnwellenmuster einzigartig? Prinzipiell ja.
  \item
    Sind aktive Gedanken anders als unbewusste? Aktive an Kaffee denken
    vs jemand stellt eine Tasse Kaffee auf den Tisch..
  \end{itemize}
\item
  Es gibt bereits ein Forschungsprojekt zum Thema (2011)
\item
  Was muss für 20.11 vorbereitet werden?
\item
  Wer präsentiert? Stefan und David
\item
  Brainstorming zu Projektnamen: Brain , Wave, Lock\ldots{}
\item
  Projektname: WaveMeIn oder BrainLock
\item
  hm? hm.. hmhm.. hm?
\end{itemize}

### 2014-10-31 diary/compile_diary.sh

@@ -1,0 +1,3 @@
+#find diary/ -name \*.md -type f -exec pandoc -o {}.tex {} \;
+pandoc -o diary_bundle.tex reports/*.md
+pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_1.tex

### 2014-10-31 diary/diary_bundle.tex
minor report updates

@@
   \item
-    lustiges Video über Gehirnwellen-gesteuerte Geräte am Ende zeigen
+    Warum, Was und Wie kommt auf die Folien
+  \item
+    lustiges Video über Gehirnwellen-gesteuerte Geräte am Ende zeigen,
     wenn Zeit bleibt
@@
   \item
-    http://www.dailymail.co.uk/sciencetech/article-2095214/As-scientists-discover-translate-brainwaves-words--Could-machine-read-innermost-thoughts.html
+    mehrere Links zu Artikeln in der Linksammlung hinzugefügt
+  \item
+    Gehirnwellen können gemessen und Gedanken erkannt werden
   \item
@@
     \item
-      ist es möglich für mehrere Benutzer bei
+      um gute Ergebnisse zu erzielen wird wahrscheinlich ein Training
+      notwendig sein (Unterlagen?)
     \end{itemize}
@@
   \item
-    Es ist sinnvoll ein Masterpasswort
+    Es ist sinnvoll mit dem Gerät ein Masterpasswort zu übertragen,
+    welches dann einzelne Domänengebundene Passwörter freigeben kann
+  \item
+    Dies kann z.B. durch eine erweiterung eines im Betriebssystem
+    eingebauten Keyrings erreicht werden
   \end{itemize}
@@
   \item
-    logcost EEG-Sensor
+    lowcost EEG-Sensor
   \item
@@
   \item
-    Muss an der Stirn abgenommen werden?
+    Müssen die Gehirnströme an der Stirn bzw. an mehreren Punkten
+    abgenommen werden?
   \end{itemize}
@@
   \item
-    es kann von Menschen verwendet werden die alternativen wie
+    es kann von Menschen verwendet werden, die Alternativen wie
     Fingerabdrucksensor oder Retinascans nicht verwenden können
   \item
-    Fingerabdrücke und Retinascanns (Biometrie generell) sind nach
-    Kompromittierung permanent unsicher
+    Biometrie generell ist nach Kompromittierung permanent unsicher
+    (Körperteile lassen sich schwer tauschen\ldots{})
   \item
-    Verschmutzung des Scanners nicht tragisch
+    Verschmutzung des Scanners nicht tragisch, Handschuhe und Brille
+    sind nicht im Weg -\textgreater{} Option für spezielle Situationen
+    wie Sport, Winter oder schmutzige/staubige Umgebung, eventuell auch
+    unter Wasser (wobei, wer gibt da Passwörter ein..)?
   \end{itemize}
 \item
-  Diskusion über die Sicherheit der Verbindung
+  Diskussion über die Sicherheit der Verbindung
 
@@
 \item
-  Es gibt bereits ein Forschungsprojekt zum Thema (2011)
-\item
-  Was muss für 20.11 vorbereitet werden?
+  Es gibt bereits ein Forschungsprojekt zum Thema (2011) -\textgreater{}
+  Linksammlung
 \item
@@
 \item
-  Projektname: WaveMeIn oder BrainLock
+  Projektname: WaveMeIn
 \item

### 2014-11-07 proposal/mswp-proposal.tex
Abgabe 1

@@
 
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
+\providecommand*{\ShortTitle}{WaveMeIn}
+\providecommand*{\FullTitle}{WaveMeIn: Authentication via Brain Waves}
 
@@
 \label{sect:synopsis}
+\subsection{The Project Idea:}
+WaveMeIn is a research project to create a new type of secure login mechanism based on brain waves.
 
-\note{
+At the time of this proposal the most used ways for authentication are manually typed passwords or biometric authentication methods.
+However all of the previous methods have some security problems or are simply not user-friendly.
+Typed passwords are easy to spy out simply by looking at the keyboard of the user or the traces of the fingers on touch displays.
+In the case of biometric authentication, there are for example face recognition, iris or fingerprint scans.
+Face recognition software can easily be tricked by face masks or photographs and moreover depends on good light conditions, the quality of the images of the web camera and other factors.
+Fingerprint and iris scans are the most secure options of the authentication methods mentioned before.
+However they also have many disadvantages.
+Iris scans are not practical since the hardware required cannot easily be integrated into small devices and it is not user-friendly to require the user to place his eye very close to the scanner every time he/she wants to unlock a device.
+Fingerprint sensors are known to fail to recognize the fingerprint correctly quite often and it is also a not very user-friendly authentication method for handicapped people that may not reach the sensor or may not have any fingers at all.
+
+Brain waves are a secure and user-friendly alternative authentication method.
+The idea is to create a small device, called Wavy, that can be worn at the ear of the user in the same style as bluetooth headsets are already worn for communication today.
+The Wavy measures the brain waves near the ear in case a login is required by a client device that is connected via bluetooth.
+It listens for a brain wave pattern that was previously trained by the user as a password.
+If the correct pattern was detected by the Wavy it transmits a OK signal back to the client device.
+
+\subsection{Why should somebody care?}
+Nowadays people are forced to type their passwords in public places which is a security risk and also not a very efficient way for authentication.
+Especially when typing in password on small devices such as mobile phones this authentication method is also very error prone due to the small keyboard interfaces.
+On the one side people are lazy and do not want to remember and enter long and complicated passwords, but on the other side they are also concerned about the security of their data and their privacy.
+So the users are in need of a more secure and easier way of authentication.
+
+\subsection{Who are the beneficiaries of the results?}
+Basically everybody can benefit from the WaveMeIn project since it is usable in the daily life.
+Especially for handicapped people it is a new and more easy to use option to log into their devices.
+Also it grants a higher level of security than existing authentication methods so it is also well suited for environments where higher security is needed, such as access authentication is modern research labs and government or military facilities.
+
+For our product to succeed, we need to invest into research in the area of brain wave detection and analysis.
+This investment can improve our understanding of this topic.
+After a commercial success, we have to enhance our product.
+This means we have to invest further into brain wave research.
+On the other side, we can make our world more secure.
+It makes hacking of accounts and password fraud more complicated.
+
+\subsection{Problem classification}
+The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning.
+In the field of Neuroscience it touches the areas of not invasive brain computer interfaces and neural oscillation.
+Since detecting and reliably identifying brain waves at the location near the ears is still technically immature the project can be seen as basic research in this area.
+The following research questions have to be answered before a prototype can be developed.
 \begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
+	\item Detecting braves at the ears
+	\item Recognize brain wave patters
+	\item Distinguish correct patterns from random signals
+	\item Distinguish brain waves from different users
 \end{itemize}
-\end{itemize}
-}
+On the other hand if we take the Wavy into account, which should be the resulting product, this project is also an applied research project.
+It further touches the fields of computer security and privacy.
 

### 2014-11-07 proposal/mswp-proposal.tex
Abgabe 1

@@
 \label{sect:synopsis}
-\subsection{The Project Idea:}
-WaveMeIn is a research project to create a new type of secure login mechanism based on brain waves.
+\subsection{Project Idea}
+WaveMeIn is a research project to create a new type of secure login mechanism. It consists of a small device worn by the user at the ear which authenticates the user based on brain waves.
 
+\subsection{Why do we need it?}
 At the time of this proposal the most used ways for authentication are manually typed passwords or biometric authentication methods.
@@
 
+\subsection{How does it work?}
 Brain waves are a secure and user-friendly alternative authentication method.

### 2014-11-07 diary/diary_bundle.tex
minor report changes

@@
 \item
-  \textbf{Zeit:} 12:10-
+  \textbf{Zeit:} 12:10-14:40
 \item
@@
 \end{itemize}
+
+\section{Submission 1: Synopsis - Fr 31 Okt
+2014}\label{submission-1-synopsis---fr-31-okt-2014}
+
+\begin{itemize}
+\item
+  \textbf{Anwesende:} Stefan, Alma, Thomas, David
+\item
+  \textbf{Abwesend:} Amra
+\item
+  \textbf{Ort:} Freihaus, Arbeitsraum gelb
+\item
+  \textbf{Zeit:} 12:15-13:50
+\item
+  \textbf{Tagesordnung:} Submission 1: Synopsis
+\item
+  Was muss in der Abgabe enthalten sein?
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    WHY, WHAT, HOW und RESULTS Beschreibung
+  \item
+    Eine ausgebaute Version der Präsentation
+  \end{itemize}
+\item
+  Das Headset wir ab jetzt Wavy, der Computer oder Handy als Device
+  genannt
+\item
+  Wie soll das Pairen funktionieren?
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    Ähnlich wie Bluetooth
+  \end{itemize}
+\item
+  Soll das Brainpattern am Wavy oder dem Device ausgewertet werden?
+
+  \begin{itemize}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    Wenn am Wavy, dann kann man das Pattern einmal trainieren und auf
+    mehreren Devices verwenden, solange diese gepairt wurden.
+  \item
+    Das trainieren ist wahrscheinlich aufwändiger und kann am Device
+    erfolgen (Patterns werden übertragen und ausgewertet)
+  \end{itemize}
+\item
+  Wie funktioniert das Entsperren?
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \itemsep1pt\parskip0pt\parsep0pt
+  \item
+    Wavy ist im passive Modus, dh. hört auf Aktivierungssignale
+  \item
+    Benutzer aktiviert das Device (z.B. den Entsperrbutton am Handy um
+    den Screen für die Mustereingabe anzuzeigen)
+  \item
+    Device sucht bekannte Wavys und sendet diesen ein
+    Authentication-Request
+  \item
+    Wavy wird aktiviert, messt Hirnwellen, wertet diese aus und
+    antwortet ggf. positiv (auch negativ??) an das Device
+  \item
+    Bei positiver Antwort wird das Device entsperrt
+  \end{enumerate}
+\item
+  Ob sich jemand wirklich dieses Protokoll ansehen wird?
+\item
+  Nächstes Treffen findet am Freitag, 14.11 wieder am selben Ort statt
+\end{itemize}

### 2014-11-21 proposal/mswp-proposal.tex

@@
 \label{sect:intro}
+WaveMeIn is a research project to show the potential of brain waves a new method of electronic authentication via a small wearable device.
+Its aim is to investigate the usage of rain waves to replace passwords or other authentication methods.
+Therefore the properties of brain waves regarding uniqueness and reliability have to be explored.
+The project shall demonstrate via a small prototype that the recognition is possible without large sensors on top of the users head.
+A requirement for such a method of authentication clearly exists as demonstrated by the following use case:
+
+Assume ...
 
@@
 \label{sect:goals}
+The following sections will provide an overview over the research questions and hardware questions associated with the project.
+\subsection{Research questions}
+\begin{itemize}
+	\item How can be brain waves be detected by the a small device at a single location?
+	\item How reliable is the detection of individual brain waves of the same person?
+	\item How reliable is unique identification of the brain waves of different persons?
+	\item Is it possible to detect brain waves at other body locations than the head?
+\end{itemize}
 
-\note{
+\subsection{Hardware Design}
 \begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
+	\item How can the required hardware be minimized to be small and practical (Wavy Device)?
+	\item Are the existing sensors for measuring rain waves good enough for the projects requirements?
+\end{itemize}
+
+\subsection{Expected Results}
 \begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
+	\item Successful research on the identification of brain waves.
+	\item Algorithms to reliably identify brain wave patterns.
+	\item Creation of a small prototype device capable of reading brain waves.
 \end{itemize}
-	\item Which results should be achieved with the project?
+
+At the end of the project it should be clear if:
 \begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
+	\item The detection of brain waves is possible at different locations of the body.
+	\item The same thought produces a repeatable and reliable brainwave pattern. (Reliability)
+	\item Different people have different patterns when thinking the same thought. (Uniqueness)
+	\item Brain waves can be used as authentication method.
+	\item The necessary can e integrated into a small device.
 \end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
+
+\subsection{Non-Goals}
+\begin{itemize}
+	\item No mind reading device
+	\item No client software (just the brain wave research, hardware and interface)
+	\item No design or usability study (just a prototype that works and is small enough)
+	\item No end-user/consumer product (just a prototype)
 \end{itemize}
-}
 

### 2014-11-21 proposal/mswp-proposal.tex
braaaaaaains yummi

@@
 The project shall demonstrate via a small prototype that the recognition is possible without large sensors on top of the users head.
-A requirement for such a method of authentication clearly exists as demonstrated by the following use case:
+A requirement for such a method of authentication clearly exists as demonstrated by the following use cases:
 
-Assume ...
+\subsection{Use Case 1}{
+Assume a user needs to log on to a device (e.g.
+a notebook) that contains sensitive information in public.
+Typing in the password is not an option as it can easily be monitored by another person.
+Fingerprints are also not a good alternative as they can easily be taken from any surface the user touched and be copied onto synthetic materials to deceive the fingerprint reader.
+Brain waves are (as of current knowledge) unique for each person even if two people are having exactly the same thought.
+If the intruder does not know the precise brain wave pattern of the user's pass phrase/thought it is impossible to duplicate.}
+
+\subsection{Use Case 2}{
+Assume an average user wants to unlock his/her smart phone in a crowded area such as the subway.
+Nowadays this is done by entering a pin or drawing a pattern on the screen.
+A person with the intention to steal a users phone just needs to observe its victim while entering the pass code or pattern.
+Afterwards it is easy to unlock the phone and steal the victims personal data or cause large costs while using it for phone calls and mobile data.
+Locking the phone via a brain wave authentication mechanism may not prevent the theft but the costs arising from the phone being used afterwards.}
+
+\subsection{Current Authentication Methods}
+In theory brain waves will be rank among the most secure authentication methods, probably being the most secure one if the research proves successful.
+The particular brain wave of a user required to unlock a device can not be easily be obtained other than strapping the user to a chair and forcing him/her to think his/her pass thought.
+Other authentication methods are password, drawing patterns, fingerprint, iris scan, voice recognition.
+Passwords and pattern drawing are the least secure ones as the user can be observed while typing or drawing without much effort.
+Fingerprint, iris scan, voice recognition may require more technical or social effort to obtain, but in the end all of them are features of a person that are always visible for the outside world and therefore copyable with more or less effort.
+
+\subsection{Unresolved Problems and Opportunities}
+The unknown factor of this research project is that no research has been done on measuring brain waves at other locations (e.g.
+the ears) of the body except directly at the users head.
+Additionally it is unknown if the brain waves of are person are distinctive enough to distinguish a pass thought of a user from other thoughts and if the brain waves of different users while thinking the same thought are distinctive enough.
+
+At the time of writing this proposal there exists no device that is capable of the features mentioned above as well as being small enough to be worn as an accessory.
+Therefore this is an important area of research with practical future applications.
+
+\subsection{Domain Specific Terms}
+\begin{itemize}
+	\item Brain waves: 
+	\item 
+\end{itemize}
 

### 2014-11-24 proposal/mswp-proposal.tex
move sections into separate files, notes on scientific relevance

@@
 % --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-\subsection{Project Idea}
-WaveMeIn is a research project to create a new type of secure login mechanism. It consists of a small device worn by the user at the ear which authenticates the user based on brain waves.
-
-\subsection{Why do we need it?}
-At the time of this proposal the most used ways for authentication are manually typed passwords or biometric authentication methods.
-However all of the previous methods have some security problems or are simply not user-friendly.
-Typed passwords are easy to spy out simply by looking at the keyboard of the user or the traces of the fingers on touch displays.
-In the case of biometric authentication, there are for example face recognition, iris or fingerprint scans.
-Face recognition software can easily be tricked by face masks or photographs and moreover depends on good light conditions, the quality of the images of the web camera and other factors.
-Fingerprint and iris scans are the most secure options of the authentication methods mentioned before.
-However they also have many disadvantages.
-Iris scans are not practical since the hardware required cannot easily be integrated into small devices and it is not user-friendly to require the user to place his eye very close to the scanner every time he/she wants to unlock a device.
-Fingerprint sensors are known to fail to recognize the fingerprint correctly quite often and it is also a not very user-friendly authentication method for handicapped people that may not reach the sensor or may not have any fingers at all.
-
-\subsection{How does it work?}
-Brain waves are a secure and user-friendly alternative authentication method.
-The idea is to create a small device, called Wavy, that can be worn at the ear of the user in the same style as bluetooth headsets are already worn for communication today.
-The Wavy measures the brain waves near the ear in case a login is required by a client device that is connected via bluetooth.
-It listens for a brain wave pattern that was previously trained by the user as a password.
-If the correct pattern was detected by the Wavy it transmits a OK signal back to the client device.
-
-\subsection{Why should somebody care?}
-Nowadays people are forced to type their passwords in public places which is a security risk and also not a very efficient way for authentication.
-Especially when typing in password on small devices such as mobile phones this authentication method is also very error prone due to the small keyboard interfaces.
-On the one side people are lazy and do not want to remember and enter long and complicated passwords, but on the other side they are also concerned about the security of their data and their privacy.
-So the users are in need of a more secure and easier way of authentication.
-
-\subsection{Who are the beneficiaries of the results?}
-Basically everybody can benefit from the WaveMeIn project since it is usable in the daily life.
-Especially for handicapped people it is a new and more easy to use option to log into their devices.
-Also it grants a higher level of security than existing authentication methods so it is also well suited for environments where higher security is needed, such as access authentication is modern research labs and government or military facilities.
-
-For our product to succeed, we need to invest into research in the area of brain wave detection and analysis.
-This investment can improve our understanding of this topic.
-After a commercial success, we have to enhance our product.
-This means we have to invest further into brain wave research.
-On the other side, we can make our world more secure.
-It makes hacking of accounts and password fraud more complicated.
-
-\subsection{Problem classification}
-The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning.
-In the field of Neuroscience it touches the areas of not invasive brain computer interfaces and neural oscillation.
-Since detecting and reliably identifying brain waves at the location near the ears is still technically immature the project can be seen as basic research in this area.
-The following research questions have to be answered before a prototype can be developed.
-\begin{itemize}
-	\item Detecting braves at the ears
-	\item Recognize brain wave patters
-	\item Distinguish correct patterns from random signals
-	\item Distinguish brain waves from different users
-\end{itemize}
-On the other hand if we take the Wavy into account, which should be the resulting product, this project is also an applied research project.
-It further touches the fields of computer security and privacy.
-
+\input{sections/synopsis}
 % --------------------------------------------------------------
@@
 A requirement for such a method of authentication clearly exists as demonstrated by the following use cases:
+
+\subsection{Current Situation}
 
@@
 
-\subsection{Domain Specific Terms}
+\subsection{Terms}
 \begin{itemize}
+	\item 
 	\item Brain waves:
@@
 \end{itemize}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
 
@@
 % --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
+\input{sections/relevance}
+% --------------------------------------------------------------
 
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{State of the art / current knowledge}
@@
 % --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
+\input{sections/method}
 % --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
+\input{sections/workplan}
 % --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
+\input{sections/timeplan}
 % --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
+\input{sections/team}
 % --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
+\input{sections/costs}
 % --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
+\input{sections/implication-risk}
 % --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
+\input{sections/ethics-security}
 % --------------------------------------------------------------

### 2014-11-24 proposal/sections/costs.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Costs}
@@
 }
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/ethics-security.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Ethical considerations \& security issues}
@@
 }
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/implication-risk.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Expected implications and risks}
@@
 }
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/method.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Method}
@@
 }
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/relevance.tex
move sections into separate files, notes on scientific relevance

\section{Scientific relevance and innovative aspects}
\label{sect:relevance}

\note{
\begin{itemize}
\item {\em Length: 1-2 pages}
\item Why is the project scientifically interesting?
\item Did others point out that this is an open question?
\item What are the innovative aspects that make it interesting?
\item How could the project break new ground scientifically?
\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
\end{itemize}
}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
WaveMeIn is an important step in the development of brain-computer interfaces used on a daily basis. Given the huge possibility, many private corporations as well as research institutes initiated promising projects ***hier ein paar links aus state of the art...***

The use case of brain wave controlled interaction in WaveMeIn is still kept simple as it is used only to unlock and no further commands have to be recognized. On success, following projects can base more sophisticated ways of interaction on results of WaveMeIns research.

* project touches several fields like computer science, neuroscience or even industrial design.
* many projects in this area, often in cooperation with industry <- results could change interaction completely

 general:
* foundation to more sophisticated applications of brainwave based interaction concepts
* a (pretty) simple way to show brainwave based interaction and pushes the public and scientific  interest

field of neuroscience/computerscience:
* do patterns differ depending on kontext (noisy environment, sleepy person or under influence of drugs/medication/zustandsverändernde substanzen)
* are the differences calculable, means, can a normalized pattern be calculated from an unknown kontext?

field of neuroscience and hw -design!!?!??!:
* optimize brain-computer-interface: smaller in size - possible to find all required bw patterns on one spot on the head?, position-independent - is it possible to find pattern in an area where its feasible to use a everyday weareble as bci? (eg not forehead but ear)

computer science ->pattern recognition (in welchen wissenschaftlichen bereichen relevant/beheimatet?) 
* pattern recognition already good, probabbly different approach needed if sensors do not record the same amount or as detailed as with bigger bci..? ->sinnvoll?!
* security: since the password in this case, the bw pattern is never shown in public, the transport from the sensor to the computer is the primary attack vector (!?) -> tight protocoll and gadget encryption is a must

The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning. In the field of Neuroscience it touches the areas of not invasive brain computer interfaces and neural oscillation. Since detecting and reliably identifying brain waves at the location near the ears is still technically immature the project can be seen as basic research in this area. The following research questions have to be answered before a prototype can be developed.
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%http://www.technologyreview.com/news/513861/samsung-demos-a-tablet-controlled-by-your-brain/
\em{http://www.essp.utdallas.edu/uploads/Main/Publications/WH2013\_Demo\_Paper.pdf}

### 2014-11-24 proposal/sections/synopsis.tex
move sections into separate files, notes on scientific relevance

\section{Synopsis}
\label{sect:synopsis}
\subsection{Project Idea}
WaveMeIn is a research project to create a new type of secure login mechanism. It consists of a small device worn by the user at the ear which authenticates the user based on brain waves.

\subsection{Why do we need it?}
At the time of this proposal the most used ways for authentication are manually typed passwords or biometric authentication methods. However all of the previous methods have some security problems or are simply not user-friendly. Typed passwords are easy to spy out simply by looking at the keyboard of the user or the traces of the fingers on touch displays. In the case of biometric authentication, there are for example face recognition, iris or fingerprint scans. Face recognition software can easily be tricked by face masks or photographs and moreover depends on good light conditions, the quality of the images of the web camera and other factors. Fingerprint and iris scans are the most secure options of the authentication methods mentioned before. However they also have many disadvantages. Iris scans are not practical since the hardware required cannot easily be integrated into small devices and it is not user-friendly to require the user to place his eye very close to the scanner every time he/she wants to unlock a device. Fingerprint sensors are known to fail to recognize the fingerprint correctly quite often and it is also a not very user-friendly authentication method for handicapped people that may not reach the sensor or may not have any fingers at all.

\subsection{How does it work?}
Brain waves are a secure and user-friendly alternative authentication method. The idea is to create a small device, called Wavy, that can be worn at the ear of the user in the same style as bluetooth headsets are already worn for communication today. The Wavy measures the brain waves near the ear in case a login is required by a client device that is connected via bluetooth. It listens for a brain wave pattern that was previously trained by the user as a password. If the correct pattern was detected by the Wavy it transmits a OK signal back to the client device.

\subsection{Why should somebody care?}
Nowadays people are forced to type their passwords in public places which is a security risk and also not a very efficient way for authentication. Especially when typing in password on small devices such as mobile phones this authentication method is also very error prone due to the small keyboard interfaces. On the one side people are lazy and do not want to remember and enter long and complicated passwords, but on the other side they are also concerned about the security of their data and their privacy. So the users are in need of a more secure and easier way of authentication.

\subsection{Who are the beneficiaries of the results?}
Basically everybody can benefit from the WaveMeIn project since it is usable in the daily life. Especially for handicapped people it is a new and more easy to use option to log into their devices. Also it grants a higher level of security than existing authentication methods so it is also well suited for environments where higher security is needed, such as access authentication is modern research labs and government or military facilities.

For our product to succeed, we need to invest into research in the area of brain wave detection and analysis. This investment can improve our understanding of this topic. After a commercial success, we have to enhance our product. This means we have to invest further into brain wave research. On the other side, we can make our world more secure. It makes hacking of accounts and password fraud more complicated.

\subsection{Problem classification}
The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning. In the field of Neuroscience it touches the areas of not invasive brain computer interfaces and neural oscillation. Since detecting and reliably identifying brain waves at the location near the ears is still technically immature the project can be seen as basic research in this area. The following research questions have to be answered before a prototype can be developed.
\begin{itemize}
	\item Detecting braves at the ears
	\item Recognize brain wave patters
	\item Distinguish correct patterns from random signals
	\item Distinguish brain waves from different users
\end{itemize}
On the other hand if we take the Wavy into account, which should be the resulting product, this project is also an applied research project. It further touches the fields of computer security and privacy.

### 2014-11-24 proposal/sections/team.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Human resources / team}
@@
 }
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/timeplan.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Detailed description of the workpackages}
-\label{sect:workplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-4 pages}
-\item Structuring the project into self-contained parts.
-\item Additional verbal descriptions.
-\item Work packages
-    \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-    \end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Time plan (Gantt chart)}
@@
 }
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-24 proposal/sections/workplan.tex
move sections into separate files, notes on scientific relevance

--- 2014-10-28 mswp-proposal.tex
@@
-\documentclass[a4paper,11pt]{article}
-
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage[utf8]{inputenc}
-\usepackage{tabularx}
-\usepackage{hyperref}
-\usepackage{color}
-\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
-% \usepackage{mathptmx} % Times New Roman
-
-\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
-\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
-\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\footskip}{15mm}
-
-\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}
-
-% --------------------------------------------------------------
-
-\providecommand*{\ShortTitle}{$<$Short title$>$}
-\providecommand*{\FullTitle}{$<$Full title$>$}
-
-% --------------------------------------------------------------
-
-\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
-{\textbf{\sffamily\Large \FullTitle}}
-\vspace{1cm}}
-
-\author{
-{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
-Group: $<$GID$>$\bigskip \\
-$<$Author1$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author1@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author2$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author2@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author3$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author3@tuwien.ac.at}{author1@tuwien.ac.at}}\\
-$<$Author4$>$ \\ {\small $<$Matrikelnr.$>$, $<$Kennzahl$>$, \href{mailto:author4@tuwien.ac.at}{author1@tuwien.ac.at}}\\ 
-\vspace{4cm}
-}
-
-\begin{document}
-
-\begin{titlepage}
-\maketitle
-
-\end{titlepage}
-
-% --------------------------------------------------------------
-
-\thispagestyle{empty}
-\tableofcontents
-\pagebreak
-
-\setcounter{page}{1}
-
-% --------------------------------------------------------------
-
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
-\pagebreak
-
-% --------------------------------------------------------------
-\section{Synopsis}
-\label{sect:synopsis}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item Concise presentation of the scientific proposal, with particular attention to the ground-breaking nature of the research project, which will allow to assess the feasibility of the outlined scientific approach.
-	\item WHY would you like to do WHAT, HOW would you like to do it and what are the expected RESULTS?
-	\item Short description of project idea.
-	\item Short description of problem / situation.
-	\item What makes it an interesting endeavor?
-	\item Why should somebody care?
-	\item Very short cost-benefit description (what are potential benefits that make them worth the costs of the project?).
-	\item Who are the beneficiaries of the results?
-	\item Problem classification (to which scientific area does the project belong?; is it basic or applied research?)
-	\item Constraints
-	\begin{itemize}
-		\item     The project must have an extent that demands to handle it as a project including project organization.  (No 2 person pieces; no 150 person-year efforts)
-		\item     Software development should be part of the project. (But not its only content or scientific contribution.)
-		\item     The project should be interesting and meaningful.
-		\item     {\em Your creativity is called for!!}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 2-3 pages}
-	\item {\bf Why?}
-	\item Introduction
-	\item Context
-	\item What is the current situation?
-	\item What is the open/unresolved problem or opportunity?
-	\item Why is it a problem?
-	\item What is unknown?
-	\item What could be improved?
-	\item Explanation of fundamental terms and basic definitions.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-
-\note{
-\begin{itemize}
-	\item {\em Length: 1-2 pages}
-	\item What is the goal of the project?
-	\item Research questions
-	\begin{itemize}
-		\item 	    What are the hypotheses that are to be investigated?
-		\item 	    Main hypothesis \& sub hypotheses
-	\end{itemize}
-	\item Which results should be achieved with the project?
-	\begin{itemize}
-	    	\item 	   What will be known afterwards that is not known now?
-		\item	    What will be created that does not exist now?
-	\end{itemize}
-	\item Non-goals (What will not be part of the project? What will not be done?)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Scientific relevance and innovative aspects}
-\label{sect:relevance}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{State of the art / current knowledge}
-\label{sect:star}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Method}
-\label{sect:method}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
-
-% --------------------------------------------------------------
 \section{Detailed description of the workpackages}
@@
 }
-
-% --------------------------------------------------------------
-\section{Time plan (Gantt chart)}
-\label{sect:timeplan}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Human resources / team}
-\label{sect:team}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
-	\begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
-		\begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
-		\end{itemize}
-	\end{itemize}
-\item Cooperations
-	\begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Costs}
-\label{sect:costs}
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Expected implications and risks}
-\label{sect:implication-risk}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
-	\begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
-	\end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
-	\begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
-	\end{itemize}
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-\section{Ethical considerations \& security issues}
-\label{sect:ethics-security}
-
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
-
-% --------------------------------------------------------------
-% APPENDIX
-\begin{appendix}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% References
-\phantomsection
-\addcontentsline{toc}{section}{References}
-
-\bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
-
-\pagebreak
-
-% --------------------------------------------------------------
-% Abbreviations
-\section*{Abbreviations}
- \addcontentsline{toc}{section}{Abbreviations}
-
- \begin{description}
-  \item[MSWP] Management von Software Projekten
-  \item[WP] Work Package
- \end{description}
-
-\end{appendix}
-
-\end{document}

### 2014-11-26 proposal/mswp-proposal.tex
möööööööööööp

--- 2014-11-21 proposal/mswp-proposal.tex
@@
 \usepackage{color}
+\usepackage{float}
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
@@
 
+\subsection{Current Situation}
+
 \subsection{Use Case 1}{
@@
 \label{sect:star}
+\subsection{What results and approaches have already been presented in this or related areas?}
+\chapter{1. Berkeley researches replace passwords by measuring brainwaves as a biometric identifier}
+\href{http://people.ischool.berkeley.edu/~chuang/pubs/usec13.pdf}
+The US Berkeley represents an approach, which turns the brain activity of an user into a biometric identifier.
+To do this, the Berkeley researchers use a 100 dollar commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode.
+This electrode is placed on the users forehead, over the brain's left frontal lobe.
+The electrode measures the users brainwaves and transmits them via a Bluetooth link to a Device.
+According to the Berkeley researchers this system has an error rate of below 1 percent.
+
+To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kind of tasks.
+For example, the participants where asked to just sit and focus on breathing in and out, imagine moving their finger up and down and listen for an audio tone.
+The participants where also asked to focus on a personalized secret, such as singing a song of their choice.
+During this tasks, the participants where wearing the EEG which measured their brain waves.
+As result came out, that not only the measured brainwaves of personalized secrets, but also measured brainwaves of simple tasks, like sitting and focusing on breathing, provided a pattern which makes it possible to authenticate an identity.
+\begin{figure}[H]
+	\centering
+    \includegraphics[width=0.5\textwidth]{john_chuang}
+    \caption{Professor John Chuang with the Neurosky MindSet brainwave sensor.}
+\end{figure}
+\chapter{2. Using Brain Waves as New Biometric Feature for Authenticating a Computer User in Real Time}
+\href{http://www.cscjournals.org/manuscript/Journals/IJBB/volume7/Issue1/IJBB-211.pdf}
+Using brain waves as a biometric Feature for Authenticating was proposed by Kusuma Mohanchandra, Lingaraju G M, Prashanth Kambli \& Vinay in the International Journal of Biometrics and Bioinformat.
+In this work it has been proved that the brain-wave pattern of every individual is  unique and the signals captured through the  EEG can be used for biometric authentication.
+This research team used an EEG EPOC headset with 14 channels to measure the brain waves.
+The collected data, containing the fusion of delta, alpha, theta, beta and gama brain waves, was merged with the aim to create a way to authenticate the user.
+
+There are three basic forms of authentication; something-you-have, something-you-know, and something-you-are.
+\begin{itemize}
+\item Something-you-have can be objects like a key or passport and people have to be very careful not to loose the object or get it stolen.
+\item Something-you-know is based on secret knowledge like passwords or PIN codes and the secret must never be written down, forgotten, or told to others.
+\item Something-you-are involves person specific features like fingerprints, voice, face, and gait.
+Authentication based on such features is called biometric authentication.
+Brain wave based authentication is a combination of something-you-know and something-you-are when the person involved has to think about something specific, but it can also be just something-you-are when the brain waves are used directly as a biometric.
+\end{itemize}
+
+The most important part of any authentication system is that true identities (clients) are verified and that false identities (impostors) are rejected.
+In a password system the password is either right or wrong, but with biometric authentication there is an uncertainty involved because the equipment that  measure the biometric feature rarely provide exactly the same data twice.
+The reason is that external parameters like finger placement, head rotation, facial hair, location etc are present.
+The challenge is to overcome these problems in such a way that even two slightly different sets of data can be verified to originate from the same person.
+There is usually a threshold that decide how different
+two different sets of data is allowed to be before they are rejected, and as a consequence there is a chance that some clients are falsely rejected and some impostors are falsely verified.
+Biometric authentication therefore introduce two error rates; False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system.
+As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.
+
+\chapter{3. Google Glass hack allows brainwave control}
+
+\subsection{Relation to the international scientific work in the field (international status of the research)}
+The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).
+It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves.
+Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.
+
+This was also explored by Benedicenti L., Koles’ Z., Mahovsky J., \& Paranjape, R.
+in the year 2001.
+This researchers used EEG directly as a biometric and their work showed some promising results on this field.
+
+EEG based person authentication was first proposed by Marcel and Millan in 2007 in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”.
+They proposed the use of  Power Spectral Density as the feature, and a statistical framework based on Gaussian Mixture  Models (GMM) and Maximum A Posteriori Model (MAP) Adaptation on speaker and face  authentication.
+The potential of their method is shown by simulations using strict train/test protocols and results.
+
+In 2001 Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”.
+To prove the connection between a person's EEG and genetically specific information, this researchers did experiments with the EEG data of healthy individuals.
+The proposed method has had a success rate of 80 percent to 100 percent showing that the EEG holds genetic information, which can be used for person identification.
+
+Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in 2008 (“Two-stage biometric authentication method using thought activity brain waves”).
+Their results show that the combination of two-stage authentication with EEG features has good potential as a biometric as it is highly resistant to fraud.
+
+\subsection{Description and critical discussion of related scientific work}
+The approaches that have been discussed above have succeeded to prove the connection between a person's DNA and her/his brain waves.
+The approaches clearly show that every single person possesses unique brain waves, which can be used to identify a person.
+However, a large disadvantage of all of this approaches is that they require the user to wear a big headset with an electrode going across his/her forehead.
+None of this approaches has succeeded in measuring brain waves from different body parts in order to enable a smaller headset and therefore make it easier for the user to use it in their everyday life.
 
@@
 \bibliographystyle{apalike}
-\bibliography{mswp2013-proposal}
+\bibliography{biblography}
 

### 2014-11-26 proposal/mswp-proposal.tex
add general information about brain waves

@@
 \subsection{Relation to the international scientific work in the field (international status of the research)}
+The brain consists of billions of brain cells called neurons.
+This neurons have to communicate with each other.
+For this communication the neurons use electricity called brain waves.
+This communication is producing a lot of brain waves, which can be detected using sensible sensors such as an EEG.
+The first person, who confirm the existence of brain waves and perform the first tests was Hans Berger.
+There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling.
+
+\begin{itemize}
+ \item Gamma (27 Hz and up)
+ \item Beta (12 Hz - 27 Hz)
+ \item Alpha (8 Hz - 12 Hz)
+ \item Theta (3 Hz - 8 Hz)
+ \item Delta (0.2 Hz - 3 Hz)
+\end{itemize}
+
 The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).

### 2014-11-26 proposal/sections/relevance.tex
scientific relevance / diary - still wip

@@
 % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-WaveMeIn is an important step in the development of brain-computer interfaces used on a daily basis.
+WaveMeIn can be seen as an important step in the development of brain-computer interfaces used on a daily basis.
 Given the huge possibility, many private corporations as well as research institutes initiated promising projects ***hier ein paar links aus state of the art...***
 
-The use case of brain wave controlled interaction in WaveMeIn is still kept simple as it is used only to unlock and no further commands have to be recognized.
+The use case of brain wave controlled interaction in WaveMeIn is still kept simple, as it is used only to unlock a device and no further commands have to be recognized.
 On success, following projects can base more sophisticated ways of interaction on results of WaveMeIns research.
 
-* project touches several fields like computer science, neuroscience or even industrial design.
-* many projects in this area, often in cooperation with industry <- results could change interaction completely
+In the field of neuroscience the main question will probably be, what kind of brain waves produce recognizable patterns of the same imagination.
+Another question is, under what circumstances do brain wave scans look similar.
+Does the pattern change if the context of the person changes, like a noisy environment, strong emotions or the effect of drugs?
+The link to pattern matching in computer science would be, how to match the original password-pattern, recorded in a probably neutral state and the input-pattern within a shifted context.
+This leads to the question, if a brain wave pattern can be normalized without knowledge of the specific context.
 
- general:
-* foundation to more sophisticated applications of brainwave based interaction concepts
-* a (pretty) simple way to show brainwave based interaction and pushes the public and scientific  interest
+In conjunction with electrical engineering the BCI itself should be revised.
+The goal is to shrink the scanner to a minimum so it does not disturb while wearing it for many hours in public.
+There not only size matters but the position of the scanner should be as flexible as possible.
+That said, a scanner with the proportions of a Bluetooth headset seems appropriate but not feasible at the moment.
+One task is to raise the level of detail of the scanners and in the same time to suppress undesired noise.
+It is still unclear what areas of the head are viable to work with an even improved non invasive BCI.
 
-field of neuroscience/computerscience:
-* do patterns differ depending on kontext (noisy environment, sleepy person or under influence of drugs/medication/zustandsverändernde substanzen)
-* are the differences calculable, means, can a normalized pattern be calculated from an unknown kontext?
+Since WaveMeIn is not only meant simplify the unlock mechanism, but to raise the security, that will be an important task in the area oc computer science.
 
-field of neuroscience and hw -design!!?!??!:
-* optimize brain-computer-interface: smaller in size - possible to find all required bw patterns on one spot on the head?, position-independent - is it possible to find pattern in an area where its feasible to use a everyday weareble as bci?
-(eg not forehead but ear)
-
-computer science ->pattern recognition (in welchen wissenschaftlichen bereichen relevant/beheimatet?) 
-* pattern recognition already good, probabbly different approach needed if sensors do not record the same amount or as detailed as with bigger bci..? ->sinnvoll?!
 * security: since the password in this case, the bw pattern is never shown in public, the transport from the sensor to the computer is the primary attack vector (!?) -> tight protocoll and gadget encryption is a must
 
+????????????
 The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning.

### 2014-11-26 proposal/mswp-proposal.tex

--- 2014-11-24 proposal/mswp-proposal.tex
@@
 \label{sect:star}
+\subsection{What results and approaches have already been presented in this or related areas?}
+\chapter{1. Berkeley researches replace passwords by measuring brainwaves as a biometric identifier}
+\href{http://people.ischool.berkeley.edu/~chuang/pubs/usec13.pdf}
+The US Berkeley represents an approach, which turns the brain activity of an user into a biometric identifier.
+To do this, the Berkeley researchers use a 100 dollar commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode.
+This electrode is placed on the users forehead, over the brain's left frontal lobe.
+The electrode measures the users brainwaves and transmits them via a Bluetooth link to a Device.
+According to the Berkeley researchers this system has an error rate of below 1 percent.
+
+To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kind of tasks.
+For example, the participants where asked to just sit and focus on breathing in and out, imagine moving their finger up and down and listen for an audio tone.
+The participants where also asked to focus on a personalized secret, such as singing a song of their choice.
+During this tasks, the participants where wearing the EEG which measured their brain waves.
+As result came out, that not only the measured brainwaves of personalized secrets, but also measured brainwaves of simple tasks, like sitting and focusing on breathing, provided a pattern which makes it possible to authenticate an identity.
+\begin{figure}[H]
+	\centering
+    \includegraphics[width=0.5\textwidth]{john_chuang}
+    \caption{Professor John Chuang with the Neurosky MindSet brainwave sensor.}
+\end{figure}
+\chapter{2. Using Brain Waves as New Biometric Feature for Authenticating a Computer User in Real Time}
+\href{http://www.cscjournals.org/manuscript/Journals/IJBB/volume7/Issue1/IJBB-211.pdf}
+Using brain waves as a biometric Feature for Authenticating was proposed by Kusuma Mohanchandra, Lingaraju G M, Prashanth Kambli \& Vinay in the International Journal of Biometrics and Bioinformat.
+In this work it has been proved that the brain-wave pattern of every individual is  unique and the signals captured through the  EEG can be used for biometric authentication.
+This research team used an EEG EPOC headset with 14 channels to measure the brain waves.
+The collected data, containing the fusion of delta, alpha, theta, beta and gama brain waves, was merged with the aim to create a way to authenticate the user.
+
+There are three basic forms of authentication; something-you-have, something-you-know, and something-you-are.
+\begin{itemize}
+\item Something-you-have can be objects like a key or passport and people have to be very careful not to loose the object or get it stolen.
+\item Something-you-know is based on secret knowledge like passwords or PIN codes and the secret must never be written down, forgotten, or told to others.
+\item Something-you-are involves person specific features like fingerprints, voice, face, and gait.
+Authentication based on such features is called biometric authentication.
+Brain wave based authentication is a combination of something-you-know and something-you-are when the person involved has to think about something specific, but it can also be just something-you-are when the brain waves are used directly as a biometric.
+\end{itemize}
+
+The most important part of any authentication system is that true identities (clients) are verified and that false identities (impostors) are rejected.
+In a password system the password is either right or wrong, but with biometric authentication there is an uncertainty involved because the equipment that  measure the biometric feature rarely provide exactly the same data twice.
+The reason is that external parameters like finger placement, head rotation, facial hair, location etc are present.
+The challenge is to overcome these problems in such a way that even two slightly different sets of data can be verified to originate from the same person.
+There is usually a threshold that decide how different
+two different sets of data is allowed to be before they are rejected, and as a consequence there is a chance that some clients are falsely rejected and some impostors are falsely verified.
+Biometric authentication therefore introduce two error rates; False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system.
+As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.
+
+\chapter{3. Google Glass hack allows brainwave control}
+
+\subsection{Relation to the international scientific work in the field (international status of the research)}
+The brain consists of billions of brain cells called neurons.
+This neurons have to communicate with each other.
+For this communication the neurons use electricity called brain waves.
+This communication is producing a lot of brain waves, which can be detected using sensible sensors such as an EEG.
+The first person, who confirm the existence of brain waves and perform the first tests was Hans Berger.
+There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling.
+
+\begin{itemize}
+ \item Gamma (27 Hz and up)
+ \item Beta (12 Hz - 27 Hz)
+ \item Alpha (8 Hz - 12 Hz)
+ \item Theta (3 Hz - 8 Hz)
+ \item Delta (0.2 Hz - 3 Hz)
+\end{itemize}
+
+The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).
+It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves.
+Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.
+
+This was also explored by Benedicenti L., Koles’ Z., Mahovsky J., \& Paranjape, R.
+in the year 2001.
+This researchers used EEG directly as a biometric and their work showed some promising results on this field.
+
+EEG based person authentication was first proposed by Marcel and Millan in 2007 in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”.
+They proposed the use of  Power Spectral Density as the feature, and a statistical framework based on Gaussian Mixture  Models (GMM) and Maximum A Posteriori Model (MAP) Adaptation on speaker and face  authentication.
+The potential of their method is shown by simulations using strict train/test protocols and results.
+
+In 2001 Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”.
+To prove the connection between a person's EEG and genetically specific information, this researchers did experiments with the EEG data of healthy individuals.
+The proposed method has had a success rate of 80 percent to 100 percent showing that the EEG holds genetic information, which can be used for person identification.
+
+Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in 2008 (“Two-stage biometric authentication method using thought activity brain waves”).
+Their results show that the combination of two-stage authentication with EEG features has good potential as a biometric as it is highly resistant to fraud.
+
+\subsection{Description and critical discussion of related scientific work}
+The approaches that have been discussed above have succeeded to prove the connection between a person's DNA and her/his brain waves.
+The approaches clearly show that every single person possesses unique brain waves, which can be used to identify a person.
+However, a large disadvantage of all of this approaches is that they require the user to wear a big headset with an electrode going across his/her forehead.
+None of this approaches has succeeded in measuring brain waves from different body parts in order to enable a smaller headset and therefore make it easier for the user to use it in their everyday life.
 

### 2014-11-26 proposal/mswp-proposal.tex
add text for google glass in chapter 5.

--- 2014-11-26 proposal/mswp-proposal.tex
@@
 \chapter{3. Google Glass hack allows brainwave control}
+After Google has released his Google Glass, a company called This Place, has developed a app to control the glass over brain waves.
+But the app alone is not enough to control the glass with the power of your mind.
+Normally the Goggle Glass is controlled over voice commands or over a touchpad built into the side of the device.
+To control the device with brain waves This Place combine the Google Glass with the Neurosky MindWave headset.
+The Neurosky MindWave is an EEG-headset to detect brain waves.
+This is one of the first devices for consumers.
+Normally this headset will be used to train your brain and it will be delivered with a hand full games.
+After combining the two devices, the company begins to testing what kind of brain waves they could detect.
+For this reason they create a simple app which implements a counter starting at 0.
+When the user concentrated, the app begin counting upwards towards 50.
+When the user starts to relaxe, the counter began returning to 0.
+After this successfully test, they want to develop an app to show a real-world use.
+So they developed MindRDR.
+It allows a user to take a photo and share it over Twitter only to use his mind.
+Very important to know is that the brain waves only activate the camera app at Google Glass and take a photo.
+The settings for Twitter were set before.
+They released their software for free, in hope that other developers can adapt it for other uses.
 

### 2014-11-26 proposal/sections/relevance.tex
extends scientific relevance

@@
 
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Why is the project scientifically interesting?
-\item Did others point out that this is an open question?
-\item What are the innovative aspects that make it interesting?
-\item How could the project break new ground scientifically?
-\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
-\end{itemize}
-}
+%\note{
+%\begin{itemize}
+%\item {\em Length: 1-2 pages}
+%\item Why is the project scientifically interesting?
+%\item Did others point out that this is an open question?
+%\item What are the innovative aspects that make it interesting?
+%\item How could the project break new ground scientifically?
+%\item To what extent are the objectives ambitious and beyond the state of the art (e.g. novel concepts and approaches or development across disciplines)?
+%\end{itemize}
+%}
 % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 WaveMeIn can be seen as an important step in the development of brain-computer interfaces used on a daily basis.
-Given the huge possibility, many private corporations as well as research institutes initiated promising projects ***hier ein paar links aus state of the art...***
+Given the huge possibility, many private corporations as well as research institutes initiated promising projects.
+To get a quick overview see Section \ref{sect:star}.
 
+\subsection{Simple step into real life}
 The use case of brain wave controlled interaction in WaveMeIn is still kept simple, as it is used only to unlock a device and no further commands have to be recognized.
 On success, following projects can base more sophisticated ways of interaction on results of WaveMeIns research.
+Even if this project covers a lot of ground work as well, the focus is to create a product usable on a daily basis.
+Therefore it goes a little further then most other project in this field, as they concentrate on mostly one specific aspect.
 
+\subsection{Bain wave recognition}
 In the field of neuroscience the main question will probably be, what kind of brain waves produce recognizable patterns of the same imagination.
@@
 
+\subsection{BCI improvement}
 In conjunction with electrical engineering the BCI itself should be revised.
@@
 
-Since WaveMeIn is not only meant simplify the unlock mechanism, but to raise the security, that will be an important task in the area oc computer science.
+\subsection{Device security.. todo}
+Since WaveMeIn is not only meant to simplify the unlock mechanism, but to raise the security of the procedure as well, this will be an important task in the area of computer science.
+The password itself is a interpretation of a specific thought and therefore never conventionally visible as maybe a fingerprint or a typed password.
+But still, it will be scanned, processed and the interpretation itself or at least a answer will be sent to the device that is waiting to get unlocked.
+The most vulnerable moment is during transport of the data.
+The security requirement should be comparable to wireless networks or Bluetooth connections and is therefore a well researched area already.
 
-* security: since the password in this case, the bw pattern is never shown in public, the transport from the sensor to the computer is the primary attack vector (!?) -> tight protocoll and gadget encryption is a must
+%* security: since the password in this case, the bw pattern is never shown in public, the transport from the sensor to the computer is the primary attack vector (!?) -> tight protocoll and gadget encryption is a must
 
-????????????
-The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning.
+%The task of detecting brain waves it tightly connected to the research areas of Neuroscience, Pattern Recognition and Machine Learning.
 In the field of Neuroscience it touches the areas of not invasive brain computer interfaces and neural oscillation.
@@
 %http://www.technologyreview.com/news/513861/samsung-demos-a-tablet-controlled-by-your-brain/
-\em{http://www.essp.utdallas.edu/uploads/Main/Publications/WH2013\_Demo\_Paper.pdf}
+%\em{http://www.essp.utdallas.edu/uploads/Main/Publications/WH2013\_Demo\_Paper.pdf}

### 2014-11-27 proposal/mswp-proposal.tex
Added sections for state-of-the-art, introduction and project-goals

--- 2014-11-26 proposal/mswp-proposal.tex
@@
 
-\note{
-\textbf{Formal constraints}
-\begin{itemize}
-\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
-\item    Fontsize: 11pt
-\item     Single line spacing
-\item     Margins: 2.5cm side and top/bottom
-\item     \fbox{Language: ENGLISH}
-\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
-\end{itemize}
-\textbf{Available templates}
-\begin{itemize}
-\item     Proposal (mswp-proposal.tex)
-\item     Costs (costs.xls, costs.ods)
-\end{itemize}
-\textbf{Supplemental material}
-\begin{itemize}
-\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
-\item     Ethical issues form (ethical-issues.rtf)
-\end{itemize}
-}
+%\note{
+%\textbf{Formal constraints}
+%\begin{itemize}
+%\item	  Font: Times New Roman oder Computer Modern (\LaTeX default)
+%\item    Fontsize: 11pt
+%\item     Single line spacing
+%\item     Margins: 2.5cm side and top/bottom
+%\item     \fbox{Language: ENGLISH}
+%\item    The proposal template should be filled incrementally. I.e., at the end there should be a full project proposal in a single PDF file.
+%\end{itemize}
+%\textbf{Available templates}
+%\begin{itemize}
+%\item     Proposal (mswp-proposal.tex)
+%\item     Costs (costs.xls, costs.ods)
+%\end{itemize}
+%\textbf{Supplemental material}
+%\begin{itemize}
+%\item     FWF salary scheme (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
+%\item     Travel cost regulation (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html})
+%\item     Ethical issues form (ethical-issues.rtf)
+%\end{itemize}
+%}
 \pagebreak
@@
 % --------------------------------------------------------------
-\section{Introduction and problem description}
-\label{sect:intro}
-WaveMeIn is a research project to show the potential of brain waves a new method of electronic authentication via a small wearable device.
-Its aim is to investigate the usage of rain waves to replace passwords or other authentication methods.
-Therefore the properties of brain waves regarding uniqueness and reliability have to be explored.
-The project shall demonstrate via a small prototype that the recognition is possible without large sensors on top of the users head.
-A requirement for such a method of authentication clearly exists as demonstrated by the following use cases:
-
-\subsection{Current Situation}
-
-\subsection{Use Case 1}{
-Assume a user needs to log on to a device (e.g.
-a notebook) that contains sensitive information in public.
-Typing in the password is not an option as it can easily be monitored by another person.
-Fingerprints are also not a good alternative as they can easily be taken from any surface the user touched and be copied onto synthetic materials to deceive the fingerprint reader.
-Brain waves are (as of current knowledge) unique for each person even if two people are having exactly the same thought.
-If the intruder does not know the precise brain wave pattern of the user's pass phrase/thought it is impossible to duplicate.}
-
-\subsection{Use Case 2}{
-Assume an average user wants to unlock his/her smart phone in a crowded area such as the subway.
-Nowadays this is done by entering a pin or drawing a pattern on the screen.
-A person with the intention to steal a users phone just needs to observe its victim while entering the pass code or pattern.
-Afterwards it is easy to unlock the phone and steal the victims personal data or cause large costs while using it for phone calls and mobile data.
-Locking the phone via a brain wave authentication mechanism may not prevent the theft but the costs arising from the phone being used afterwards.}
-
-\subsection{Current Authentication Methods}
-In theory brain waves will be rank among the most secure authentication methods, probably being the most secure one if the research proves successful.
-The particular brain wave of a user required to unlock a device can not be easily be obtained other than strapping the user to a chair and forcing him/her to think his/her pass thought.
-Other authentication methods are password, drawing patterns, fingerprint, iris scan, voice recognition.
-Passwords and pattern drawing are the least secure ones as the user can be observed while typing or drawing without much effort.
-Fingerprint, iris scan, voice recognition may require more technical or social effort to obtain, but in the end all of them are features of a person that are always visible for the outside world and therefore copyable with more or less effort.
-
-\subsection{Unresolved Problems and Opportunities}
-The unknown factor of this research project is that no research has been done on measuring brain waves at other locations (e.g.
-the ears) of the body except directly at the users head.
-Additionally it is unknown if the brain waves of are person are distinctive enough to distinguish a pass thought of a user from other thoughts and if the brain waves of different users while thinking the same thought are distinctive enough.
-
-At the time of writing this proposal there exists no device that is capable of the features mentioned above as well as being small enough to be worn as an accessory.
-Therefore this is an important area of research with practical future applications.
-
-\subsection{Terms}
-\begin{itemize}
-	\item 
-	\item Brain waves:
-	\item
-\end{itemize}
-
+\input{sections/introduction}
 % --------------------------------------------------------------
-\section{Project goals and deliverables}
-\label{sect:goals}
-The following sections will provide an overview over the research questions and hardware questions associated with the project.
-\subsection{Research questions}
-\begin{itemize}
-	\item How can be brain waves be detected by the a small device at a single location?
-	\item How reliable is the detection of individual brain waves of the same person?
-	\item How reliable is unique identification of the brain waves of different persons?
-	\item Is it possible to detect brain waves at other body locations than the head?
-\end{itemize}
-
-\subsection{Hardware Design}
-\begin{itemize}
-	\item How can the required hardware be minimized to be small and practical (Wavy Device)?
-	\item Are the existing sensors for measuring rain waves good enough for the projects requirements?
-\end{itemize}
-
-\subsection{Expected Results}
-\begin{itemize}
-	\item Successful research on the identification of brain waves.
-	\item Algorithms to reliably identify brain wave patterns.
-	\item Creation of a small prototype device capable of reading brain waves.
-\end{itemize}
-
-At the end of the project it should be clear if:
-\begin{itemize}
-	\item The detection of brain waves is possible at different locations of the body.
-	\item The same thought produces a repeatable and reliable brainwave pattern. (Reliability)
-	\item Different people have different patterns when thinking the same thought. (Uniqueness)
-	\item Brain waves can be used as authentication method.
-	\item The necessary can e integrated into a small device.
-\end{itemize}
-
-\subsection{Non-Goals}
-\begin{itemize}
-	\item No mind reading device
-	\item No client software (just the brain wave research, hardware and interface)
-	\item No design or usability study (just a prototype that works and is small enough)
-	\item No end-user/consumer product (just a prototype)
-\end{itemize}
-
+\input{sections/project-goals}
 % --------------------------------------------------------------
@@
 % --------------------------------------------------------------
-
-\section{State of the art / current knowledge}
-\label{sect:star}
-\subsection{What results and approaches have already been presented in this or related areas?}
-\chapter{1. Berkeley researches replace passwords by measuring brainwaves as a biometric identifier}
-\href{http://people.ischool.berkeley.edu/~chuang/pubs/usec13.pdf}
-The US Berkeley represents an approach, which turns the brain activity of an user into a biometric identifier.
-To do this, the Berkeley researchers use a 100 dollar commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode.
-This electrode is placed on the users forehead, over the brain's left frontal lobe.
-The electrode measures the users brainwaves and transmits them via a Bluetooth link to a Device.
-According to the Berkeley researchers this system has an error rate of below 1 percent.
-
-To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kind of tasks.
-For example, the participants where asked to just sit and focus on breathing in and out, imagine moving their finger up and down and listen for an audio tone.
-The participants where also asked to focus on a personalized secret, such as singing a song of their choice.
-During this tasks, the participants where wearing the EEG which measured their brain waves.
-As result came out, that not only the measured brainwaves of personalized secrets, but also measured brainwaves of simple tasks, like sitting and focusing on breathing, provided a pattern which makes it possible to authenticate an identity.
-\begin{figure}[H]
-	\centering
-    \includegraphics[width=0.5\textwidth]{john_chuang}
-    \caption{Professor John Chuang with the Neurosky MindSet brainwave sensor.}
-\end{figure}
-\chapter{2. Using Brain Waves as New Biometric Feature for Authenticating a Computer User in Real Time}
-\href{http://www.cscjournals.org/manuscript/Journals/IJBB/volume7/Issue1/IJBB-211.pdf}
-Using brain waves as a biometric Feature for Authenticating was proposed by Kusuma Mohanchandra, Lingaraju G M, Prashanth Kambli \& Vinay in the International Journal of Biometrics and Bioinformat.
-In this work it has been proved that the brain-wave pattern of every individual is  unique and the signals captured through the  EEG can be used for biometric authentication.
-This research team used an EEG EPOC headset with 14 channels to measure the brain waves.
-The collected data, containing the fusion of delta, alpha, theta, beta and gama brain waves, was merged with the aim to create a way to authenticate the user.
-
-There are three basic forms of authentication; something-you-have, something-you-know, and something-you-are.
-\begin{itemize}
-\item Something-you-have can be objects like a key or passport and people have to be very careful not to loose the object or get it stolen.
-\item Something-you-know is based on secret knowledge like passwords or PIN codes and the secret must never be written down, forgotten, or told to others.
-\item Something-you-are involves person specific features like fingerprints, voice, face, and gait.
-Authentication based on such features is called biometric authentication.
-Brain wave based authentication is a combination of something-you-know and something-you-are when the person involved has to think about something specific, but it can also be just something-you-are when the brain waves are used directly as a biometric.
-\end{itemize}
-
-The most important part of any authentication system is that true identities (clients) are verified and that false identities (impostors) are rejected.
-In a password system the password is either right or wrong, but with biometric authentication there is an uncertainty involved because the equipment that  measure the biometric feature rarely provide exactly the same data twice.
-The reason is that external parameters like finger placement, head rotation, facial hair, location etc are present.
-The challenge is to overcome these problems in such a way that even two slightly different sets of data can be verified to originate from the same person.
-There is usually a threshold that decide how different
-two different sets of data is allowed to be before they are rejected, and as a consequence there is a chance that some clients are falsely rejected and some impostors are falsely verified.
-Biometric authentication therefore introduce two error rates; False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system.
-As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.
-
-\chapter{3. Google Glass hack allows brainwave control}
-
-\subsection{Relation to the international scientific work in the field (international status of the research)}
-The brain consists of billions of brain cells called neurons.
-This neurons have to communicate with each other.
-For this communication the neurons use electricity called brain waves.
-This communication is producing a lot of brain waves, which can be detected using sensible sensors such as an EEG.
-The first person, who confirm the existence of brain waves and perform the first tests was Hans Berger.
-There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling.
-
-\begin{itemize}
- \item Gamma (27 Hz and up)
- \item Beta (12 Hz - 27 Hz)
- \item Alpha (8 Hz - 12 Hz)
- \item Theta (3 Hz - 8 Hz)
- \item Delta (0.2 Hz - 3 Hz)
-\end{itemize}
-
-The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).
-It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves.
-Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.
-
-This was also explored by Benedicenti L., Koles’ Z., Mahovsky J., \& Paranjape, R.
-in the year 2001.
-This researchers used EEG directly as a biometric and their work showed some promising results on this field.
-
-EEG based person authentication was first proposed by Marcel and Millan in 2007 in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”.
-They proposed the use of  Power Spectral Density as the feature, and a statistical framework based on Gaussian Mixture  Models (GMM) and Maximum A Posteriori Model (MAP) Adaptation on speaker and face  authentication.
-The potential of their method is shown by simulations using strict train/test protocols and results.
-
-In 2001 Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”.
-To prove the connection between a person's EEG and genetically specific information, this researchers did experiments with the EEG data of healthy individuals.
-The proposed method has had a success rate of 80 percent to 100 percent showing that the EEG holds genetic information, which can be used for person identification.
-
-Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in 2008 (“Two-stage biometric authentication method using thought activity brain waves”).
-Their results show that the combination of two-stage authentication with EEG features has good potential as a biometric as it is highly resistant to fraud.
-
-\subsection{Description and critical discussion of related scientific work}
-The approaches that have been discussed above have succeeded to prove the connection between a person's DNA and her/his brain waves.
-The approaches clearly show that every single person possesses unique brain waves, which can be used to identify a person.
-However, a large disadvantage of all of this approaches is that they require the user to wear a big headset with an electrode going across his/her forehead.
-None of this approaches has succeeded in measuring brain waves from different body parts in order to enable a smaller headset and therefore make it easier for the user to use it in their everyday life.
-
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item What results and approaches have already been presented in this or related areas?
-\item Relation to the international scientific work in the field (international status of the research)
-\item Description and critical discussion of related scientific work
-\end{itemize}
-}
-
+%\note{
+%\begin{itemize}
+%\item {\em Length: 2-5 pages}
+%\item What results and approaches have already been presented in this or related areas?
+%\item Relation to the international scientific work in the field (international status of the research)
+%\item Description and critical discussion of related scientific work
+%\end{itemize}
+%}
+\input{sections/state-of-the-art}
 % --------------------------------------------------------------

### 2014-11-27 proposal/sections/introduction.tex
Added sections for state-of-the-art, introduction and project-goals

\section{Introduction and problem description}
\label{sect:intro}
WaveMeIn is a research project to show the potential of brain waves a new method of electronic authentication via a small wearable device. Its aim is to investigate the usage of rain waves to replace passwords or other authentication methods. Therefore the properties of brain waves regarding uniqueness and reliability have to be explored. The project shall demonstrate via a small prototype that the recognition is possible without large sensors on top of the users head. A requirement for such a method of authentication clearly exists as demonstrated by the following use cases:

\subsection{Use Case 1}{
Assume a user needs to log on to a device (e.g. a notebook) that contains sensitive information in public. Typing in the password is not an option as it can easily be monitored by another person. Fingerprints are also not a good alternative as they can easily be taken from any surface the user touched and be copied onto synthetic materials to deceive the fingerprint reader. Brain waves are (as of current knowledge) unique for each person even if two people are having exactly the same thought. If the intruder does not know the precise brain wave pattern of the user's pass phrase/thought it is impossible to duplicate.}

\subsection{Use Case 2}{
Assume an average user wants to unlock his/her smart phone in a crowded area such as the subway. Nowadays this is done by entering a pin or drawing a pattern on the screen. A person with the intention to steal a users phone just needs to observe its victim while entering the pass code or pattern. Afterwards it is easy to unlock the phone and steal the victims personal data or cause large costs while using it for phone calls and mobile data. Locking the phone via a brain wave authentication mechanism may not prevent the theft but the costs arising from the phone being used afterwards.}

\subsection{Current Authentication Methods}
In theory brain waves will be rank among the most secure authentication methods, probably being the most secure one if the research proves successful. The particular brain wave of a user required to unlock a device can not be easily be obtained other than strapping the user to a chair and forcing him/her to think his/her pass thought. Other authentication methods are password, drawing patterns, fingerprint, iris scan, voice recognition. Passwords and pattern drawing are the least secure ones as the user can be observed while typing or drawing without much effort. Fingerprint, iris scan, voice recognition may require more technical or social effort to obtain, but in the end all of them are features of a person that are always visible for the outside world and therefore copyable with more or less effort.

\subsection{Unresolved Problems and Opportunities}
The unknown factor of this research project is that no research has been done on measuring brain waves at other locations (e.g. the ears) of the body except directly at the users head. Additionally it is unknown if the brain waves of are person are distinctive enough to distinguish a pass thought of a user from other thoughts and if the brain waves of different users while thinking the same thought are distinctive enough.

At the time of writing this proposal there exists no device that is capable of the features mentioned above as well as being small enough to be worn as an accessory. Therefore this is an important area of research with practical future applications.
%
%\subsection{Terms}
%\begin{itemize}
%	\item 
%	\item Brain waves:
%	\item
%\end{itemize}

### 2014-11-27 proposal/sections/project-goals.tex
Added sections for state-of-the-art, introduction and project-goals

\section{Project goals and deliverables}
\label{sect:goals}
The following sections will provide an overview over the research questions and hardware questions associated with the project.
\subsection{Research questions}
\begin{itemize}
	\item How can be brain waves be detected by the a small device at a single location?
	\item How reliable is the detection of individual brain waves of the same person?
	\item How reliable is unique identification of the brain waves of different persons?
	\item Is it possible to detect brain waves at other body locations than the head?
\end{itemize}

\subsection{Hardware Design}
\begin{itemize}
	\item How can the required hardware be minimized to be small and practical (Wavy Device)?
	\item Are the existing sensors for measuring rain waves good enough for the projects requirements?
\end{itemize}

\subsection{Expected Results}
\begin{itemize}
	\item Successful research on the identification of brain waves.
	\item Algorithms to reliably identify brain wave patterns.
	\item Creation of a small prototype device capable of reading brain waves.
\end{itemize}

At the end of the project it should be clear if:
\begin{itemize}
	\item The detection of brain waves is possible at different locations of the body.
	\item The same thought produces a repeatable and reliable brainwave pattern. (Reliability)
	\item Different people have different patterns when thinking the same thought. (Uniqueness)
	\item Brain waves can be used as authentication method.
	\item The necessary can e integrated into a small device.
\end{itemize}

\subsection{Non-Goals}
\begin{itemize}
	\item No mind reading device
	\item No client software (just the brain wave research, hardware and interface)
	\item No design or usability study (just a prototype that works and is small enough)
	\item No end-user/consumer product (just a prototype)
\end{itemize}

### 2014-11-27 proposal/sections/state-of-the-art.tex
Added sections for state-of-the-art, introduction and project-goals

\section{State of the art / current knowledge}
\label{sect:star}
\subsection{What results and approaches have already been presented in this or related areas?}
\subsubsection{Berkeley researches replace passwords by measuring brainwaves as a biometric identifier}
\href{http://people.ischool.berkeley.edu/~chuang/pubs/usec13.pdf}
The US Berkeley represents an approach, which turns the brain activity of an user into a biometric identifier. To do this, the Berkeley researchers use a 100 dollar commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode. This electrode is placed on the users forehead, over the brain's left frontal lobe. The electrode measures the users brainwaves and transmits them via a Bluetooth link to a Device. According to the Berkeley researchers this system has an error rate of below 1 percent.

To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kind of tasks. For example, the participants where asked to just sit and focus on breathing in and out, imagine moving their finger up and down and listen for an audio tone. The participants where also asked to focus on a personalized secret, such as singing a song of their choice. During this tasks, the participants where wearing the EEG which measured their brain waves. As result came out, that not only the measured brainwaves of personalized secrets, but also measured brainwaves of simple tasks, like sitting and focusing on breathing, provided a pattern which makes it possible to authenticate an identity.
\begin{figure}[H]
	\centering
    \includegraphics[width=0.5\textwidth]{john_chuang}
    \caption{Professor John Chuang with the Neurosky MindSet brainwave sensor.}
\end{figure}
\subsubsection{Using Brain Waves as New Biometric Feature for Authenticating a Computer User in Real Time}
\href{http://www.cscjournals.org/manuscript/Journals/IJBB/volume7/Issue1/IJBB-211.pdf}
Using brain waves as a biometric feature for authenticating was proposed by Kusuma Mohanchandra, Lingaraju G M, Prashanth Kambli \& Vinay in the International Journal of Biometrics and Bioinformat. In this work it has been proved that the brain-wave pattern of every individual is  unique and the signals captured through the  EEG can be used for biometric authentication. This research team used an EEG EPOC headset with 14 channels to measure the brain waves. The collected data, containing the fusion of delta, alpha, theta, beta and gamma brain waves, was merged with the aim to create a way to authenticate the user.

There are three basic forms of authentication; something-you-have, something-you-know, and something-you-are.
\begin{itemize}
\item Something-you-have can be objects like a key or passport and people have to be very careful not to loose the object or get it stolen.
\item Something-you-know is based on secret knowledge like passwords or PIN codes and the secret must never be written down, forgotten, or told to others.
\item Something-you-are involves person specific features like fingerprints, voice, face, and gait. Authentication based on such features is called biometric authentication. Brain wave based authentication is a combination of something-you-know and something-you-are when the person involved has to think about something specific, but it can also be just something-you-are when the brain waves are used directly as a biometric.
\end{itemize}

The most important part of any authentication system is that true identities (clients) are verified and that false identities (impostors) are rejected. In a password system the password is either right or wrong, but with biometric authentication there is an uncertainty involved because the equipment that  measure the biometric feature rarely provide exactly the same data twice. The reason is that external parameters like finger placement, head rotation, facial hair, location etc are present. The challenge is to overcome these problems in such a way that even two slightly different sets of data can be verified to originate from the same person. There is usually a threshold that decide how different
two different sets of data is allowed to be before they are rejected, and as a consequence there is a chance that some clients are falsely rejected and some impostors are falsely verified.
Biometric authentication therefore introduce two error rates; False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system. As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.

\subsubsection{Google Glass hack allows brainwave control}
After Google has released his Google Glass, a company called This Place, has developed a app to control the glass over brain waves. But the app alone is not enough to control the glass with the power of your mind. Normally the Goggle Glass is controlled over voice commands or over a touchpad built into the side of the device. To control the device with brain waves This Place combine the Google Glass with the Neurosky MindWave headset. The Neurosky MindWave is an EEG-headset to detect brain waves. This is one of the first devices for consumers. Normally this headset will be used to train your brain and it will be delivered with a hand full games. After combining the two devices, the company begins to testing what kind of brain waves they could detect. For this reason they create a simple app which implements a counter starting at 0. When the user concentrated, the app begin counting upwards towards 50. When the user starts to relaxe, the counter began returning to 0. After this successfully test, they want to develop an app to show a real-world use. So they developed MindRDR. It allows a user to take a photo and share it over Twitter only to use his mind. Very important to know is that the brain waves only activate the camera app at Google Glass and take a photo. The settings for Twitter were set before. They released their software for free, in hope that other developers can adapt it for other uses.

\subsection{Relation to the international scientific work in the field (international status of the research)}
The brain consists of billions of brain cells called neurons. This neurons have to communicate with each other. For this communication the neurons use electricity called brain waves. This communication is producing a lot of brain waves, which can be detected using sensible sensors such as an EEG. The first person, who confirm the existence of brain waves and perform the first tests was Hans Berger. There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling.

\begin{itemize}
 \item Gamma (27 Hz and up)
 \item Beta (12 Hz - 27 Hz)
 \item Alpha (8 Hz - 12 Hz)
 \item Theta (3 Hz - 8 Hz)
 \item Delta (0.2 Hz - 3 Hz)
\end{itemize}

The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA). It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves. Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.

This was also explored by Benedicenti L., Koles’ Z., Mahovsky J., \& Paranjape, R. in the year 2001. This researchers used EEG directly as a biometric and their work showed some promising results on this field.

EEG based person authentication was first proposed by Marcel and Millan in 2007 in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”. They proposed the use of  Power Spectral Density as the feature, and a statistical framework based on Gaussian Mixture  Models (GMM) and Maximum A Posteriori Model (MAP) Adaptation on speaker and face  authentication. The potential of their method is shown by simulations using strict train/test protocols and results.

In 2001 Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”. To prove the connection between a person's EEG and genetically specific information, this researchers did experiments with the EEG data of healthy individuals. The proposed method has had a success rate of 80 percent to 100 percent showing that the EEG holds genetic information, which can be used for person identification.

Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in 2008 (“Two-stage biometric authentication method using thought activity brain waves”). Their results show that the combination of two-stage authentication with EEG features has good potential as a biometric as it is highly resistant to fraud.

\subsection{Description and critical discussion of related scientific work}
The approaches that have been discussed above have succeeded to prove the connection between a person's DNA and her/his brain waves. The approaches clearly show that every single person possesses unique brain waves, which can be used to identify a person. However, a large disadvantage of all of this approaches is that they require the user to wear a big headset with an electrode going across his/her forehead. None of this approaches has succeeded in measuring brain waves from different body parts in order to enable a smaller headset and therefore make it easier for the user to use it in their everyday life.

### 2014-11-27 proposal/sections/introduction.tex
Fixed references

@@
 \label{sect:intro}
-WaveMeIn is a research project to show the potential of brain waves a new method of electronic authentication via a small wearable device.
-Its aim is to investigate the usage of rain waves to replace passwords or other authentication methods.
+WaveMeIn is a research project to show the potential of brain waves, a new method of electronic authentication via a small wearable device.
+Its aim is to investigate the usage of brain waves to replace passwords or other authentication methods.
 Therefore the properties of brain waves regarding uniqueness and reliability have to be explored.
@@
 In theory brain waves will be rank among the most secure authentication methods, probably being the most secure one if the research proves successful.
-The particular brain wave of a user required to unlock a device can not be easily be obtained other than strapping the user to a chair and forcing him/her to think his/her pass thought.
+The particular brain wave of a user required to unlock a device can not be obtained easily other than strapping the user to a chair and forcing him/her to think his/her pass thought.
 Other authentication methods are password, drawing patterns, fingerprint, iris scan, voice recognition.

### 2014-11-27 proposal/sections/project-goals.tex
Fixed references

@@
 \begin{itemize}
-	\item How can be brain waves be detected by the a small device at a single location?
+	\item How can brain waves be detected by a small device at a single location?
 	\item How reliable is the detection of individual brain waves of the same person?
@@
 	\item How can the required hardware be minimized to be small and practical (Wavy Device)?
-	\item Are the existing sensors for measuring rain waves good enough for the projects requirements?
+	\item Are the existing sensors for measuring brain waves good enough for the projects requirements?
 \end{itemize}
@@
 	\item Brain waves can be used as authentication method.
-	\item The necessary can e integrated into a small device.
+	\item The necessary can be integrated into a small device.
 \end{itemize}

### 2014-11-27 proposal/sections/relevance.tex
Fixed references

@@
 
-\subsection{Bain wave recognition}
+\subsection{Brain wave recognition}
 In the field of neuroscience the main question will probably be, what kind of brain waves produce recognizable patterns of the same imagination.
@@
 
-\subsection{Device security.. todo}
+\subsection{Device security}
 Since WaveMeIn is not only meant to simplify the unlock mechanism, but to raise the security of the procedure as well, this will be an important task in the area of computer science.
-The password itself is a interpretation of a specific thought and therefore never conventionally visible as maybe a fingerprint or a typed password.
+The password itself is an interpretation of a specific thought and therefore never conventionally visible as maybe a fingerprint or a typed password.
 But still, it will be scanned, processed and the interpretation itself or at least a answer will be sent to the device that is waiting to get unlocked.

### 2014-11-27 proposal/sections/state-of-the-art.tex
Fixed references

@@
 The US Berkeley represents an approach, which turns the brain activity of an user into a biometric identifier.
-To do this, the Berkeley researchers use a 100 dollar commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode.
+To do this, the Berkeley researchers use a commercial EEG (electroencephalogram), which resembles a Bluetooth headseat with an electrode.
 This electrode is placed on the users forehead, over the brain's left frontal lobe.
@@
 
-To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kind of tasks.
+To ensure that the brain waves of every single person are unique and that they provide enough information to authenticate the user's identity, the Berkley researchers performed tests with participants, which included different kinds of tasks.
 For example, the participants where asked to just sit and focus on breathing in and out, imagine moving their finger up and down and listen for an audio tone.
@@
 \end{figure}
-\subsubsection{Using Brain Waves as New Biometric Feature for Authenticating a Computer User in Real Time}
+\subsubsection{Using brain waves as a new biometric feature for authenticating user in real time}
 \href{http://www.cscjournals.org/manuscript/Journals/IJBB/volume7/Issue1/IJBB-211.pdf}
@@
 
-There are three basic forms of authentication; something-you-have, something-you-know, and something-you-are.
+In \cite{kennet} three basic forms of authentication are identified: something-you-have, something-you-know, and something-you-are. According to \cite{kennet}:"
 \begin{itemize}
@@
 
-The most important part of any authentication system is that true identities (clients) are verified and that false identities (impostors) are rejected.
+The most important part of any authentication system is that true identities are verified and that false identities are rejected.
 In a password system the password is either right or wrong, but with biometric authentication there is an uncertainty involved because the equipment that  measure the biometric feature rarely provide exactly the same data twice.
@@
 two different sets of data is allowed to be before they are rejected, and as a consequence there is a chance that some clients are falsely rejected and some impostors are falsely verified.
-Biometric authentication therefore introduce two error rates; False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system.
-As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.
+Biometric authentication therefore introduce two error rates: False Non-Match Rate (FNMR), the rate at which clients are falsely rejected by the system, and False Match Rate (FMR), the rate at which impostors are falsely verified by the system.
+As such the main problem in this thesis is two compare two or more EEG signals and decide whether they are from the same person or not, and get as low FNMR and FMR as possible.''
 
 \subsubsection{Google Glass hack allows brainwave control}
-After Google has released his Google Glass, a company called This Place, has developed a app to control the glass over brain waves.
-But the app alone is not enough to control the glass with the power of your mind.
+After Google has released his Google Glass, a company called ``This Place``, has developed an app to control the glass over brain waves.
+However, the app alone is not enough to control the glass with the power of your mind.
 Normally the Goggle Glass is controlled over voice commands or over a touchpad built into the side of the device.
-To control the device with brain waves This Place combine the Google Glass with the Neurosky MindWave headset.
+To control the device with brain waves ``This Place`` combined the Google Glass with a Neurosky MindWave headset.
 The Neurosky MindWave is an EEG-headset to detect brain waves.
@@
 Normally this headset will be used to train your brain and it will be delivered with a hand full games.
-After combining the two devices, the company begins to testing what kind of brain waves they could detect.
-For this reason they create a simple app which implements a counter starting at 0.
-When the user concentrated, the app begin counting upwards towards 50.
-When the user starts to relaxe, the counter began returning to 0.
-After this successfully test, they want to develop an app to show a real-world use.
-So they developed MindRDR.
-It allows a user to take a photo and share it over Twitter only to use his mind.
-Very important to know is that the brain waves only activate the camera app at Google Glass and take a photo.
+After combining the two devices, the company begined with testing the different kinds of brain waves they could detect.
+For this reason they created a simple app which implemented a counter starting at 0.
+When the user concentrated, the app began counting upwards towards 50.
+When the user started to relax, the counter began returning to 0.
+After this successfull test, they created an app to show a real-world use, called MindRDR.
+It allows an user to take a photo and share it over Twitter only by using his mind.
+Very important to know is that the brain waves only activate the camera app at Google Glass and take a picture.
 The settings for Twitter were set before.
@@
 \subsection{Relation to the international scientific work in the field (international status of the research)}
+The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).
+It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves.
+Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.
+
 The brain consists of billions of brain cells called neurons.
@@
 This communication is producing a lot of brain waves, which can be detected using sensible sensors such as an EEG.
-The first person, who confirm the existence of brain waves and perform the first tests was Hans Berger.
-There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling.
+The first person, who confirmed the existence of brain waves and performed the first tests was Hans Berger.
+There are five different kinds of brain waves and all of them are directly connected to what a person is thinking, doing and feeling:
 
@@
 
-The very beginning of the existence of brain wave based authentication systems dates back to the 1960's when Vogel discovered a connection between a person's EEG signals and his/her genetic code(DNA).
-It was proved that every person owns unique brainwaves and it possible to identify a person through his brainwaves.
-Identical twins where shown to have the same EEG patterns in the same situations and even changes related to aging were similar.
+This was also explored by \cite{Benedicenti2001} in the year 2001. This researchers used EEG directly as a biometric and their work showed some promising results on this field.
 
-This was also explored by Benedicenti L., Koles’ Z., Mahovsky J., \& Paranjape, R.
-in the year 2001.
-This researchers used EEG directly as a biometric and their work showed some promising results on this field.
-
-EEG based person authentication was first proposed by Marcel and Millan in 2007 in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”.
+EEG based person authentication was first proposed by \cite{marcel} in “Person authentication using brainwaves (EEG) and maximum a posteriori model adaption”.
 They proposed the use of  Power Spectral Density as the feature, and a statistical framework based on Gaussian Mixture  Models (GMM) and Maximum A Posteriori Model (MAP) Adaptation on speaker and face  authentication.
@@
 
-In 2001 Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”.
+In \cite{Poulos01personidentification} Poulos, Alexandris and Evangelou performed person identification based on spectral information and presented their results in their work: “On the use of EEG features towards oersin identification via neural networks”.
 To prove the connection between a person's EEG and genetically specific information, this researchers did experiments with the EEG data of healthy individuals.
@@
 
-Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in 2008 (“Two-stage biometric authentication method using thought activity brain waves”).
+Furthermore, a novel two-stage biometric authentication method was proposed by Palaniappan in \cite{Palaniappan08}.
 Their results show that the combination of two-stage authentication with EEG features has good potential as a biometric as it is highly resistant to fraud.

### 2014-11-27 diary/10_Diary_2.tex
Added project diary 2

\documentclass[a4paper,11pt]{article}

\usepackage{graphicx}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
% \usepackage{mathptmx} % Times New Roman

\setlength{\topmargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textheight}{243.119mm} % 297mm-40mm-10mm-(11pt=3.881mm)=
\setlength{\oddsidemargin}{-0.4mm} % (1in=25.4mm)-0.4mm=25mm
\setlength{\textwidth}{160mm} % 210mm-50mm=160mm
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\footskip}{15mm}

\providecommand*{\note}[1]{\small \textcolor{RoyalBlue}{\begin{minipage}{\textwidth}{#1}\end{minipage}}}

% --------------------------------------------------------------

\providecommand*{\ShortTitle}{WaveMeIn}
\providecommand*{\FullTitle}{Projektprotokoll}
\providecommand*{\includeWiki}[1]{\wikiEnvironments \include{#1}\nowikiEnvironments}

% --------------------------------------------------------------

\title{\textbf{\sffamily\Huge \ShortTitle}\\ 
{\textbf{\sffamily\Large \FullTitle}}
\vspace{1cm}}

\author{
{\em 188.407: Management von Software Projekten} \vspace{1cm} \\
Group: 10\bigskip \\
Belk Stefan \\ {\small 0750926, 937, \href{mailto:belk.stefan@gmail.com}{belk.stefan@gmail.com}}\\
Petz Thomas \\ {\small 0601280, 937, \href{mailto:e0601280@student.tuwien.ac.at}{e0601280@student.tuwien.ac.at}}\\
Causevic Alma \\ {\small 0847805, 534, \href{mailto:alma.causevic@hotmail.com}{alma.causevic@hotmail.com}}\\ 
Causevic Amra  \\ {\small 0649241, 534, \href{mailto:amra.causevic@hotmail.com}{amra.causevic@hotmail.com}}\\ 
Seebacher David \\ {\small 0327243, 534, \href{mailto:david.seebacher@student.tuwien.ac.at}{david.seebacher@student.tuwien.ac.at}}\\
\vspace{4cm}
}

\begin{document}

\begin{titlepage}
\maketitle

\end{titlepage}

% --------------------------------------------------------------

\section{Submission 2 -Treffen Freihaus am 20.11.2014}

\begin{itemize}
\item
  \textbf{Anwesende:} Alle
\item
  \textbf{Abwesend:} Niemand
\item
  \textbf{Ort:} Freihaus, Arbeitsraum gelb
\item
  \textbf{Zeit:} 13:00 - 17:00
\item
  \textbf{Tagesordnung:} Gemeinsame Diskussion für die Kapitel 2 bis 5
\begin{itemize}
  \itemsep1pt\parskip0pt\parsep0pt
  \item Goals, Non Goals abgegrentzt
  \item Resaerch questions formuliert
  \item Arbeitsaufteilung
\end{itemize}
\end{itemize}

\section{Weiterer Verlauf: Gegenseitige Fortschrittsberichte und Hilfestellung über Emails.}

\subsection{21.11.2014 - Stefan}
Hi,

Also wir sind wie erwartet nicht ganz fertig geworden, aber haben schon einiges geschafft.
Alma und Amra bitte schaut euch Kapitel 2 an und erarbeitet Kapitel 5.

Kapitel 2)
Hab ich angefangen. Eine Seite hab ich schon geschrieben, am Ende fahlt noch ein bisschen was. Werde daran noch weiter machen. 
@Amra/Alma: Bitte drüberlesen und vielleicht noch den einen oder anderen Satz einfügen wenn euch was dazu einfällt.

Kapitel 3)
Ist fertig.

Kapitel 4) 
David hat gesagt das Kapitel macht er fertig.

Kapitel 5)
Hat der Thomas angefangen (ein paar Zeilen).
@Amra/Alma: Bitte ausarbeiten. Müssen keine 5 Seiten sein, 2-3 reichen auch. Bitte teilt euch ein wer da welche der im Kapitel gestellten Fragen bearbeitet.

LG Stefan

\subsection{24.11.2014 - David}
Hey,

hier ein kleines Update von mir. Hab zu Kapitel 4 ca ne Seite Stichworte und muss das ganze noch in einen gescheiten Text umschreiben, was ich leider gestern (und heute) nicht geschafft habe. Vielleicht habe ich heut, noch vor dem Schlafen gehen, was vorzuzeigen, realistischerweise allerdings erst morgen Abend.

lg David

\subsection{25.11.2014 - Thomas}
ja hallo,
ich weiß, dass ich spät dran bin, aber kann leider erst morgen was zu meinem teil beitragen. bin die woche ein bisschen ins strudeln gekommen ...

lg

\subsection{26.11.2014 - Amra, Alma, Stefan}
Hi,
Die Amra hat mich gebeten euch Bescheid zu geben, dass schon einiges zu kapitel 5 fertig ist.
@Thomas vor allem einiges von Berkeley und aus Papers:
- Berkeley research 
- Using brainwaves aus nem guten pdf aus cscjournals
- Das schon in dern 60ern was zum thema ga, 
- 2001 benedicenti, koles und Paranjape "the electriencephalogram as a biometric ......"
- 2008 Palaniappan rR. " two stges biometric authentication method using thought"

Die Alma überarbeitet das noch und fügt es dann am Abend ein. Schau was du noch so findest, z.B über dieses eine Startup, das GoogleGlass mit Brainwaves steuert und zu "critical discussion and description"

Stefan

\subsection{26.11.2014 - Thomas}
ok, wäre es möglich die rohfassung mal zu pushen, damit ich mal den letzten stand hab ...

\subsection{26.11.2014 - Alma}
Hi,
 anbei der derzeitige Stand des Kapitels 5.

Was noch gemacht werden sollte ist:
1 google glas
2 Description and critical discussion of related scientific work (da steht zwar was drinnen, aber ich glaube dass wir uns eher auf die nachteile des derzeitigen produktes konzentrieren sollten - größe des headsets, unpraktisch usw.

\section{Abgabe am 27.11.2014}
\begin{itemize}
  \item Gegenseitiges Lesen der Kapitel, Feedback und kleine Verbesserungen
  \item Bibtex Einträge erstellen von Alma
\end{itemize}

\section{What happened since the last meeting?}
Wir haben 15 Punkte auf die erste Abgabe bekommen!

Feedback: Die Instruktionen von Seite 3 entfernen. ...ist erledigt!

\end{document}

### 2014-11-28 diary/compile_diary.sh
add todays report, define diary file for submission 3

@@ -1,3 +1,3 @@
 #find diary/ -name \*.md -type f -exec pandoc -o {}.tex {} \;
 pandoc -o diary_bundle.tex reports/*.md
-pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_1.tex
+pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_3.tex

### 2014-12-03 diary/compile_diary.sh
Add diary 12.02 and update older ones

@@ -1,3 +1,4 @@
 #find diary/ -name \*.md -type f -exec pandoc -o {}.tex {} \;
-pandoc -o diary_bundle.tex reports/*.md
+pandoc -f markdown_phpextra -o diary_bundle.tex reports/*.md
 pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_3.tex
+rm *.out *.log *.synctex.gz *.aux

### 2014-12-09 proposal/sections/workplan.tex
add workplan from googledoc to tex.

@@
 }
+
+\subsection{Basic research}
+\begin{itemize}
+ \item goal(s): The goal of the research part is to clarify the scientific requirements, needed for the creation of the prototype.
+ \item description: This work package deals with the brain waves related part of this project.
+ \item expected results: Thought identification, confirmation of the uniqueness of brain waves, measuring brain waves from different body parts.
+ \item responsible person: Neuroscience researcher
+ \item dependencies: none
+ \item start date: 05.01.2015
+ \item end date: 05.07.2016
+\end{itemize}
+
+\subsubsection{Confirm uniqueness of a person’s brain wave pattern}
+\begin{itemize}
+ \item goal(s): Confirm assumption of persons unique brain waves.
+ \item description: Every person has an unique brain wave pattern for the same thought. This pattern can be used to identify the person.
+ \item expected results: If the research shows unexpected results and it the uniqueness can’t be proved, it will limit the security aspect of the project, but it still can be realised.
+ \item responsible person: Neuroscience researcher
+ \item dependencies: none
+ \item start date: 05.01.2015
+ \item end date: 05.07.2015
+\end{itemize}
+
+\subsubsection{Reliable thought identification}
+\begin{itemize}
+ \item goal(s): Confirmation of our assumption, that a thought of a person can be identified again as the same.
+ \item description: Research on how a specific brain wave of a person can be identified reliably.
+ \item expected results: A person can be identified again as the same.
+ \item responsible person: Neuroscience researcher
+ \item dependencies: WP 1.1
+ \item start date: 05.07.2015
+ \item end date: 05.01.2016
+\end{itemize}
+
+\subsubsection{Measuring methods of brain waves from different parts of the head}
+\begin{itemize}
+ \item goal(s): Existing studies showed theoretical approaches to gatter brain waves with in-ear or behind-ear EEGs.
+The goal is to use this approaches and define requirements for building a small device to measure brain waves.
+ \item description: In this package we attempt to define requirements for building a small device.
+ \item expected results: Prove that brain waves can be measured from different head-parts than the forehead.
+ \item responsible person: Neuroscience researcher
+ \item dependencies: WP 1.2
+ \item start date: 05.01.2016
+ \item end date: 05.07.2016
+\end{itemize}
+
+\subsubsection{Survey of state of the art for sensors and devices for brain wave measurement}
+\begin{itemize}
+ \item goal(s): Analyse the hardware components of existing brain wave measurement and gather enough information to build a small device.
+ \item description: There already exists a lot of sensors for detecting brain waves. Main part here is to look, if a device already exists or if we buy a existing one a make it smaller.
+ \item expected results: Today do probably exist decent sensors, but none exactly as needed for the planned tiny device.
+ \item responsible person: Hardware engineer
+ \item dependencies: none
+ \item start date: 05.01.2015
+ \item end date: 05.01.2016
+\end{itemize}
+
+\subsection{Hardware and software development}
+\begin{itemize}
+ \item goal(s): Develop prototype hardware and software for showing the success.
+ \item description: To show that authentication over brain waves works, we have to develop a hardware prototype.
+There different kinds of software we have to develop.
+First an algorithm to ensure fault tolerance and an interface for the communication with the device.
+ \item expected results: The developed application shows a successfully communication with the device.
+ \item responsible person: Hardware engineer
+ \item dependencies: WP 1.2
+ \item start date: 05.07.2016
+ \item end date: 05.07.2017
+\end{itemize}
+
+\subsubsection{Reviewing of existing algorithms for detecting patterns in measured brain}
+\begin{itemize}
+ \item goal(s): To identify the thought of a person.
+ \item description: It is possible that the same thought of a person can be shown different in the brain wave. For this case we need some patterns to confirm the measured brain waves.
+ \item expected results: A working algorithm with good performance.
+ \item responsible person: Computer science researcher
+ \item dependencies: none
+ \item start date: 05.07.2016
+ \item end date: 05.10.2016
+\end{itemize}
+
+\subsubsection{Develop new algorithms for detecting patterns in measured brain waves}
+\begin{itemize}
+ \item goal(s):  Develop the needed algorithms.
+ \item description: If there are no existing algorithms for detecting patterns in measured brain waves, which can be used in this project, then new ones need to be developed.
+ \item expected results: Our algorithms need to be reliable and fast.
+ \item responsible person: Computer science researcher
+ \item dependencies: WP 2.1 
+ \item start date: 05.10.2016
+ \item end date: 05.01.2017
+\end{itemize}
+
+\subsubsection{Define the communication protocol and interface of the Wavy}
+\begin{itemize}
+ \item goal(s): A software interface for the Wavy
+ \item description: Develop an software interface for other applications to communicate with the Wavy.
+This interface should operation system or other developers allow to handle their authentication over Wavy.
+ \item expected results: A easy usable and reliable software interface implement for several programing languages.
+ \item responsible person: Software engineer
+ \item dependencies: none
+ \item start date: 05.07.2016
+ \item end date: 05.07.2017
+\end{itemize}
+
+\subsubsection{Develop small sensors for brain wave measurement}
+\begin{itemize}
+ \item goal(s): Develop small sensors for brain wave measurement so they can be placed in the Wavy.
+ \item description: Because the Wavy should be build as tiny as possible small sensors are needed.
+ \item expected results: This work results in a tiny fully functional sensor for surement.
+ \item responsible person: Hardware engineer
+ \item dependencies: WP 1.4
+ \item start date: 05.07.2016
+ \item end date: 05.01.2017
+\end{itemize}
+
+\subsubsection{Develop a microcontroller for the Wavy}
+\begin{itemize}
+ \item goal(s): A microcontroller to handle and identify brain waves.
+ \item description: Develop a microcontroller thats manage the detection and identification of brain waves.
+It is working with the developed algorithm and sending it over a bluetooth interface to a chosen device.
+The communication with the controller is handled with the developed software interface.
+ \item expected results: A working microcontroller.
+ \item responsible person: Hardware engineer
+ \item dependencies: WP 2.2, WP 2.3
+ \item start date: 05.01.2017
+ \item end date: 05.07.2017
+\end{itemize}
+
+\subsection{Prototype development}
+\begin{itemize}
+ \item goal(s): A device that is similar than the device for the consumer.
+ \item description: Develop a prototype which can satisfy all requirements
+ \item expected results:
+ \begin{itemize}
+  \item Wavy prototype
+  \item Demonstration software
+  \item A study confirming the superiority of our new method
+ \end{itemize}
+ \item responsible person: Hardware engineer
+ \item dependencies: WP 2
+ \item start date: 05.07.2017
+ \item end date: 05.07.2018
+\end{itemize}
+
+\subsubsection{Create the Wavy prototype}
+\begin{itemize}
+ \item goal(s): Develop a wearable prototype.
+ \item description: The Wavy prototype doesn’t have to be the final product but it should function like specified above.
+ \item expected results: working prototype (Wavy)
+ \item responsible person: Hardware engineer
+ \item dependencies: WP 2
+ \item start date: 05.07.2017
+ \item end date: 05.03.2018
+\end{itemize}
+
+\subsubsection{Write a software prototype}
+\begin{itemize}
+ \item goal(s): Write a simple prototype including all essential features of WaveMeIn.
+ \item description: The software prototype should implement the software, which is required for using the WaveMeIn Features.
+ \item expected results: A small software to show how the Wavy device works (log in to an operation system).
+ \item responsible person: Software engineer
+ \item dependencies: WP 2
+ \item start date: 05.07.2017
+ \item end date: 05.03.2018
+\end{itemize}
+
+\subsubsection{Comparing with existing authentication methods}
+\begin{itemize}
+ \item goal(s): certificate, publicity
+ \item description: An external research comparison.
+ \item expected results: Confirm that the safety of this authentication method is superior to any other authentication method.
+ \item responsible person: Computer science researcher
+ \item dependencies: WP 2, WP 3.1
+ \item start date: 05.03.2018
+ \item end date: 05.07.2018
+\end{itemize}

### 2014-12-11 proposal/sections/method.tex

@@
 \label{sect:method}
+\subsection{How should the expected results be achieved?}
+In the first place some basic research about brain waves should be done in order to achieve the expected results and build a device for brain wave based authentication.
+The goal of the projects research part is to clarify the scientific requirements needed for the creation of the prototype, for example to confirm the uniqueness of brain waves or define how specific brain waves of a person can be identified.
+This is important since we want to assure that the user can be identified as the same one later again.
+It should be guaranteed that thoughts provide a reliable identification method, therefore the research should give answers how reliably brain waves of a person can be identified.
 
-\note{
-\begin{itemize}
-\item {\em Length: 2-5 pages}
-\item {\bf How?}
-\item How should the expected results be achieved?
-\item What method(s) will be applied? (e.g., empirical study, user-centered design, prototype implementation,...)
-\item Description of the methods.
-\item Justifications for chosen methods.
-\end{itemize}
-}
+The research part consists of two phases – a theoretical and an empirical research part.
+The purpose of the theoretical part is to analyse existing studies and use the results for the project.
+Already existing studies confirmed the possibility of authentication via brain waves, but so far no study managed to measure brain waves from different head parts than the forehead.
+The studies showed theoretical approaches to measure brain waves using EEG electrodes placed at the users forehead.
+The goal of the theoretical research part is to use this approaches and sensor minimising methods in order to define requirements for building a small device, called Wavy.
+Therefore, the goal to measure brain waves from different head or even body parts is a very important part of the research.
+The Wavy device should be as small as possible and designed in a way which allows the user to wear it in everyday life.
+The ears will be the primary focus of this research for measuring brain waves at positions less visible then the forehead.
+This will enable the Wavy device to be worn like an accessory on or behind the ear.
+
+In the empirical part there will be some practical tests with human beings, where their brain waves will be measured with an EEG in different situations.
+The goal of this research part is to confirm the uniqueness of two peoples brain waves, even if they think about the same thing.
+The approach from UC Berkeley, which John Chuang presented at the 17th International Conference on Financial Cryptography and Data Security, is a good one to use for the project.
+In this approach the participants were asked to perform seven mental tasks.
+These were divided into two categories.
+For the first group of tasks - which was the same for all participants - the subjects were asked to do simple things, for example to focus on their own breathing, imagine moving a finger, or to listen for an audio tone and respond to that by focusing on a dot on a piece of paper.
+For the second group of tasks - which participants selected and performed individually without letting others know what they were doing - the subjects were asked to select from imagining performing a repetitive motion from their favourite sport, such as swinging a golf club; singing a song of their choice; watch a series of on-screen objects and silently count those that matched a color of their choice; or think of something for 10 seconds.
+\cite{Adhikari13} With such tests the Berkeley researchers managed to ensure that brain waves provide enough information to authenticate the user's identity.
+This kind of tests will be also done with users in our empirical research part once the prototype will be constructed.
+
+\subsection{What methods will be applied?}
+The following methods will be used:
+
+\subsubsection{Theoretical study}
+The goal of the theoretical study is to gather information from existing studies on this field.
+Several different studies already investigated the possibility of using brain waves as access method.
+Their findings will be used as a start point for our research.
+Researchers in the field of neuroscience and computer science will be needed for this task which will be part of the first work packages at the beginning of the project.
+
+\subsubsection{Experiments}
+This part consists of performing practical tests with participants.
+During these tests the users will wear an EEG device and will be asked to think of some simple things as well as thinking of some personal memories or for example their favourite song.
+The experiments will define whether brain waves of every single person are unique and how reliable they can be measured from different body parts.
+The participants will have to be selected as heterogeneous group so that the experiments can be performed on people with different characteristics such as male/female, adult/children, etc.
+Again the neuroscience and computer science are the required fields of expertise for this part.
+
+\subsubsection{Prototype}
+In order to get to this project phase the first phase which includes the brain wave connected research part should be completed.
+The goal of this phase is to use the results from the research part and create the Wavy prototype and write a demo software.
+The Wavy should be as small as possible and communicates with the user's device via Bluetooth.
+It listens for a brain wave pattern that was previously trained by the user as a password.
+If the correct pattern was detected by the Wavy it transmits a OK signal back to the client device.
+A demonstration software shall be written to show how the Wavy device can be used.
+This work will require an hardware and a software engineer to work together with the researchers.
+
+\subsubsection{Scientific study}
+At the end of the project the security and usability of authenticating via the Wavy device will be compared to existing authentication methods.
+This study should generate a comprehensive report which can be published to show the advantages of this new authentication method and raise the public's awareness for it.

### 2014-12-11 proposal/sections/timeplan.tex

@@
 \label{sect:timeplan}
+As specified in section 7, the project consists of three basic work packages (each of them having several sub-packages):
+\begin{itemize}
+\item WP 1 - Basic research
+\item WP 2 - Hardware and software development
+\item WP 3 - Prototype development
+\end{itemize}
 
-\note{
+\begin{figure}[H]
+	\centering
+    \includegraphics[width=1\textwidth]{gantt_chart}
+    \caption{GANTT Chart}
+\end{figure}
+
+Each of these packages consists of another sub-packages.
+Since the research part of the project delivers the basis for the whole project and therefore is the most important part of the project, it has a planned duration of 1,5 years.
+The planned duration of the hardware and software development package is 1 year and the duration of the prototype development package 1 year.
+Also there is a scheduled time buffer of one month in each package, which can be used in potential problematic situations and in cases of time delays appearances.
+
+\subsection{Milestones}
+The planned milestones of the projects are:
 \begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Realistic estimation of schedule based on workpackages.
-\item Including milestones (not only when but also what is to be achieved for each milestone).
-\item Generation of a Gantt chart. (Including phases, milestones, buffer times, critical areas, etc.)
+\item 05.07.2015 - Confirmed uniqueness of a person’s brain wave pattern
+\item 05.07.2016 - Research part completed 
+\item 05.01.2017 - Hardware and Software for Microcontroller finished
+\item 05.06.2017 - Hardware Part Completed 
+\item 05.02.2018 - Prototype Completed 
+\item 05.07.2018 - Project Completed 
 \end{itemize}
-}
+
+\subsection{Timeplan}
+The estimation of the schedule based on work packages:
+
+\subsubsection{WP1 - Basic research}
+\begin{itemize}
+\item start date: 05.01.2015
+\item end date:  05.07.2016
+\item milestone: Research part completed
+\end{itemize}
+
+\subsubsection{WP 1.1 - Confirm uniqueness of a person’s brain wave pattern}
+\begin{itemize}
+\item start date: 05.01.2015
+\item end date: 05.07.2015
+\item milestone: Confirmed uniqueness of a person’s brain wave pattern
+\end{itemize}
+\subsubsection{WP 1.2 - Reliable thought identification}
+\begin{itemize}
+\item start date: 05.07.2015
+\item end date:  05.01.2016
+\end{itemize}
+\subsubsection{WP 1.3 - Measuring methods of brain waves from different parts of the head}
+\begin{itemize}
+\item start date: 05.01.2016
+\item end date: 05.07.2016
+\end{itemize}
+\subsubsection{WP 1.4 - Survey of state of the art for sensors and devices for brain wave measurement}
+\begin{itemize}
+\item start date: 05.01.2015
+\item end date: 05.01.2016
+\end{itemize}
+\subsubsection{WP 2 - Hardware and software development}
+\begin{itemize}
+\item start date: 05.07.2016
+\item end date:  05.07.2017
+\end{itemize}
+
+The Milestones for the work packages 2.2, 2.3 and 2.4 share the same milestone - Hardware and Software for Microcontroller finished and after this three work packages there is also planned a buffer of one month for eventual time delays.
+
+\subsubsection{WP 2.1 - Reviewing of existing algorithms for detecting patterns in measured brain}
+\begin{itemize}
+\item start date: 05.07.2016
+\item end date:  05.10.2016
+\end{itemize}
+\subsubsection{WP 2.2 - Develop new algorithms for detecting patterns in measured brain waves}
+\begin{itemize}
+\item start date: 05.10.2016
+\item end date:  05.01.2017
+\item milestone: Hardware and Software for Microcontroller finished 
+\end{itemize}
+\subsubsection{WP 2.3 - Define the communication protocol and interface of the Wavy}
+\begin{itemize}
+\item start date: 05.07.2016
+\item end date:  05.07.2017
+\item milestone: Hardware and Software for Microcontroller finished 
+\end{itemize}
+\subsubsection{WP 2.4 - Develop small sensors for brain wave measurement}
+\begin{itemize}
+\item start date: 05.07.2016
+\item end date: 05.01.2017
+\item milestone: Hardware and Software for Microcontroller finished 
+\end{itemize}
+\subsubsection{WP 2.5 - Develop a microcontroller for the Wavy}
+\begin{itemize}
+\item start date: 05.01.2017
+\item end date: 05.07.2017
+\item milestone: Hardware Part Completed
+\end{itemize}
+\subsubsection{WP 3 - Prototype development}
+\begin{itemize}
+\item start date: 05.07.2017
+\item end date: 05.07.2018
+\item milestone: Project Completed
+\end{itemize}
+\subsubsection{WP 3.1  - Create the Wavy prototype}
+\begin{itemize}
+\item start date: 05.07.2017
+\item end date: 05.03.2018
+\item milestone: Prototype Completed
+\end{itemize}
+\subsubsection{WP 3.2  - Write a software prototype}
+\begin{itemize}
+\item start date: 05.07.2017
+\item end date: 05.03.2018
+\item milestone: Prototype Completed
+\end{itemize}
+\subsubsection{WP 3.3  - Comparing with existing authentication methods}
+\begin{itemize}
+\item start date: 05.03.2018
+\item end date:  05.07.2018
+\end{itemize}
+
+\subsection{Critical Areas}
+The critical areas of this project are in the research part, because if the research parts doesn’t deliver the following results, the project will not fail but it won’t lead to the planned project goals.
+The Work package 1.1 (Confirm uniqueness of a person’s brain wave pattern) and 1.3 (Identify the methods to measure brain waves from different head parts) are the most critical parts of the project.

### 2014-12-11 proposal/sections/workplan.tex
remove note from tex.

@@
 
-\note{
-\begin{itemize}
-  \item {\em Length: 2-4 pages}
-  \item Structuring the project into self-contained parts.
-  \item Additional verbal descriptions.
-  \item Work packages
-  \begin{itemize}
-    \item title
-    \item goal(s)
-    \item description
-    \item expected results
-    \item responsible person(s)
-    \item dependencies
-  \end{itemize}
-\end{itemize}
-}
-
-\subsection{Basic research}
+\subsection*{WP 1 - Basic research}
 \begin{itemize}
@@
 
-\subsubsection{Confirm uniqueness of a person’s brain wave pattern}
+\subsubsection*{WP 1.1 - Confirm uniqueness of a person’s brain wave pattern}
 \begin{itemize}
@@
 
-\subsubsection{Reliable thought identification}
+\subsubsection*{WP 1.2 - Reliable thought identification}
 \begin{itemize}
@@
 
-\subsubsection{Measuring methods of brain waves from different parts of the head}
+\subsubsection*{WP 1.3 - Measuring methods of brain waves from different parts of the head}
 \begin{itemize}
@@
 
-\subsubsection{Survey of state of the art for sensors and devices for brain wave measurement}
+\subsubsection*{WP 1.4 - Survey of state of the art for sensors and devices for brain wave measurement}
 \begin{itemize}
@@
 
-\subsection{Hardware and software development}
+\subsection*{WP 2 - Hardware and software development}
 \begin{itemize}
@@
 
-\subsubsection{Reviewing of existing algorithms for detecting patterns in measured brain}
+\subsubsection*{WP 2.1 - Reviewing of existing algorithms for detecting patterns in measured brain}
 \begin{itemize}
@@
 
-\subsubsection{Develop new algorithms for detecting patterns in measured brain waves}
+\subsubsection*{WP 2.2 - Develop new algorithms for detecting patterns in measured brain waves}
 \begin{itemize}
@@
 
-\subsubsection{Define the communication protocol and interface of the Wavy}
+\subsubsection*{WP 2.3 - Define the communication protocol and interface of the Wavy}
 \begin{itemize}
@@
 
-\subsubsection{Develop small sensors for brain wave measurement}
+\subsubsection*{WP 2.4 - Develop small sensors for brain wave measurement}
 \begin{itemize}
@@
 
-\subsubsection{Develop a microcontroller for the Wavy}
+\subsubsection*{WP 2.5 - Develop a microcontroller for the Wavy}
 \begin{itemize}
@@
 
-\subsection{Prototype development}
+\subsection*{WP 3 - Prototype development}
 \begin{itemize}
@@
 
-\subsubsection{Create the Wavy prototype}
+\subsubsection*{WP 3.1 - Create the Wavy prototype}
 \begin{itemize}
@@
 
-\subsubsection{Write a software prototype}
+\subsubsection*{WP 3.2 - Write a software prototype}
 \begin{itemize}
@@
 
-\subsubsection{Comparing with existing authentication methods}
+\subsubsection*{WP 3.3 - Comparing with existing authentication methods}
 \begin{itemize}

### 2014-12-11 proposal/sections/method.tex
Finished method section

@@
 In the first place some basic research about brain waves should be done in order to achieve the expected results and build a device for brain wave based authentication.
-The goal of the projects research part is to clarify the scientific requirements needed for the creation of the prototype, for example to confirm the uniqueness of brain waves or define how specific brain waves of a person can be identified.
+The goal of the projects research part is to clarify the scientific requirements needed for the creation of the prototype.
+There are two main scientific questions that need to be answered.
+First, confirm the uniqueness of brain waves or define how specific brain waves of a person can be identified.
 This is important since we want to assure that the user can be identified as the same one later again.
 It should be guaranteed that thoughts provide a reliable identification method, therefore the research should give answers how reliably brain waves of a person can be identified.
+Second, besides confirming the uniqueness of a persons brain waves compared to the brain waves of others, another important factor is to confirm that the brain waves of a specific thought of a person are distinguishable from other thoughts of the same person.
+To prove our assumptions about the uniqueness of brain waves experiments will be conducted on human subjects in the empirical part of this project.
 
@@
 For the second group of tasks - which participants selected and performed individually without letting others know what they were doing - the subjects were asked to select from imagining performing a repetitive motion from their favourite sport, such as swinging a golf club; singing a song of their choice; watch a series of on-screen objects and silently count those that matched a color of their choice; or think of something for 10 seconds.
-\cite{Adhikari13} With such tests the Berkeley researchers managed to ensure that brain waves provide enough information to authenticate the user's identity.
-This kind of tests will be also done with users in our empirical research part once the prototype will be constructed.
+\cite{Adhikari13} These tests will be repeated over several weeks to verify the critical uniqueness of a person's thoughts as mentioned above a prolonged period of time.
+With such tests the Berkeley researchers managed to ensure that brain waves provide enough information to authenticate the user's identity.
+This kind of tests will be also done with users in our empirical research part.
 
@@
 The participants will have to be selected as heterogeneous group so that the experiments can be performed on people with different characteristics such as male/female, adult/children, etc.
-Again the neuroscience and computer science are the required fields of expertise for this part.
+Neuroscience and computer science are the required fields of expertise for this part.
+The results of the experiments will also determine which sensors are best suited for the future prototype.
+The human brain produces a wide range of different brain waves with different frequencies.
+Some of these frequencies are more important to thought recognition than others.
+Available EEG devices can measure delta waves (0.5 to 3 Hz), theta waves (3 to 8 Hz), alpha waves (8 to 12 Hz), beta waves (12 to 38 Hz) and gamma waves (38 to 42 Hz).
+The experimental part will narrow down the range of waves that are best suited for our goals.
 

### 2015-01-08 proposal/sections/costs.tex
Costs

@@
 \label{sect:costs}
+\subsection{Person}
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline 
+Person & Commitment & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+\hline 
+Project Manager & 100\% & 54000 & • & • & • & • \\ 
+\hline 
+Quality Manager & 75\% & 27000 & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+• & • & • & • & • & • & • \\ 
+\hline 
+\end{tabular}
 
-\note{
-\begin{itemize}
-\item {\em Length: 2-3 pages}
-\item Rough estimation of cost in form of calculation (table(s)) + descriptive text.
-\item Justification for the personnel and non-personnel costs (equipment, material, travel and other costs)
-\item An Excel template is provided as supplementary material to support budgeting.
-\item Personnel costs
-	\begin{itemize}
-	\item     Justification for the personnel to be assigned to the project (type of position(s), description of nature of work, length and extent of involvement in the project)
-	\item     The application should include all persons who will be required for the proposed project (project lead, researchers, developers, advisory board, etc.).
-The available legal categories of employment are contracts of employment for full- or part-time employees (DV) and reimbursement for work on an hourly basis (GB).
-In addition, a part-time contract of employment (DV 50\%, ``studentische Mitarbeiter'') may be requested for people who have not yet completed a Master or Diploma program (Diplom) in the relevant subject.
-	 \item    The justification of the requested personnel should contain:
-		\begin{itemize}
-		\item 	        description of type of work;
-		\item 		        extent of involvement (part-time contracts are permitted).
-		\end{itemize}
-	\item Exact numbers of employment categories can be found on the FWF Website (\href{http://www.fwf.ac.at/de/projects/personalkostensaetze.html}{http://www.fwf.ac.at/de/projects/personalkostensaetze.html})
-	\end{itemize}
-\item Equipment costs
-	\begin{itemize}
-	\item     Indicate reasons for equipment costs.
-The ``scientific equipment'' category includes instruments, system components, costs for the use of software required by the project and other durable goods provided the cost per item (including VAT) exceeds EUR 1,500.00.
-	\end{itemize}
-\item Material costs
-	\begin{itemize}
-	\item     This category encompasses consumables and smaller pieces of equipment where the cost per item is below EUR 1,500.00 including VAT.
-The calculation of requested material costs should be justified with reference to the schedule, work plan and experimental plan.
-Experience with previous projects should be taken into account.
-	\end{itemize}
-\item Travel costs
-	\begin{itemize}
-	\item     Funding may be requested for the costs of project-specific travel and accommodation, field work, expeditions, etc.
-Applicants are to provide a detailed travel (cost) plan broken down by project participant.
-For brief stays, the calculation of the travel and accommodation costs should be based on the federal regulations governing travel costs (RGV).
-The RGV rates governing Austria and abroad may be found in the FAQs on the FWF Website (\href{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}{http://www.fwf.ac.at/de/faq/reisegebuehrenvorschrift.html}).
-For longer stays an appropriate and comprehensible cost plan should be prepared.
-	\end{itemize}
-\item Other costs
-	\begin{itemize}
-	\item     Independent contracts for work and services (costs for work of clearly defined scope and content assigned to individuals, provided that this is scientifically justifiable and economical)
-    	\item     Costs that cannot be included under personnel, equipment, material or travel costs, such as:
-		\begin{itemize}
-		\item         reimbursement of costs towards or for the use of research facilities, e.g.
-of large-scale research facilities (project-specific 'equipment time').
-Applicants should obtain and submit multiple offers;
-		\item         costs for project-specific work carried out outside the applicant's research institution (e.g.
-for analysis work performed elsewhere, for interviews, for sample collection, for preparation of thin slices etc.).
-Applicants should obtain and submit multiple offers;
-		\item         honoraria for test persons;
-		\end{itemize}
-	\end{itemize}
-\end{itemize}
-}
+\subsection{Hardware}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline 
+Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+\hline 
+Commercial EEG Headsets & 3000 & • & • & • & • \\ 
+\hline 
+Medical EEG Device & 60000 & • & • & • & • \\ 
+\hline 
+Server & 4000 & • & • & • & • \\ 
+\hline 
+Workstations & Notebooks & 6400 & • & • & • & • \\ 
+\hline 
+\end{tabular}
+\subsection{Material}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline 
+Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+\hline 
+Hardware for the Prototype & 2000 & • & • & • & 2000 \\ 
+\hline 
+EEG Sensors & 500 & • & • & • & 500 \\ 
+\hline 
+• & • & • & • & • & • \\ 
+\hline 
+\end{tabular} 
+\subsection{Travel Costs}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline 
+Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+\hline 
+Conferences & 4000 & • & 4000 & • & 8000 \\ 
+\hline 
+EEG Sensors & 500 & • & • & • & 500 \\ 
+\hline 
+• & • & • & • & • & • \\ 
+\hline 
+\end{tabular} 
+\subsection{Other Costs}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline 
+Item & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+\hline 
+Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 8000 \\
+\hline 
+\end{tabular}

### 2015-01-08 proposal/sections/ethics-security.tex
Costs

@@
 }
+
+\subsection{Ethical Considerations}
+\begin{itemize}
+\item {
+At the moment there is an ongoing dispute in the public over weather or not long-term exposure to electromagnetic fields can cause cancer or changes in the brain chemistry.
+As far as we know today weak electromagnetic fields will have no effects on the human brain, but strong fields definitively do.
+The Wavy device will definitly create such a weak magnetic field (like all electric devices) and due to its location at the users ear and the intention of wearing it permanently it has the potential of harming the user if electromagnetic waves really have negative effects on the human brain.
+}
+\item {
+The Wavy device could be copied by companies trying to make a profit by constantly monitoring the emotional state of the user.
+The device would still function as an access device, but it could be misused for marketing purposes for example.
+A possible scenario could be: A company installs a wireless module into a TV set that sends a signal when an advertisment is shown.
+This signal activates an application in the user’s phone which in turn starts collecting emotional data via such a “bad” Wavy device from the user who is watching the advertisment.
+The company can then sell the gathered knowledge about the user’s response to the advertisment and his/her preferences to others such as the advertising company.
+}
+\end{itemize}
+
+\subsection{Security Issues}
+\begin{itemize}
+\item{
+At the current state of the technology mind reading is still science fiction, but basically a device such as the Wavy could be used to read all thoughts of a user.
+The only real issue that is preventing this scenario at the moment is that from a given set of brain waves there is currently no way of translating this user’s thought into a visual or textual representation for someone else to read.
+However, this may be possible at some point in the future.
+A compromised Wavy device poses a serious security risk for the user.
+}
+\item{
+The biggest issue in security is probably the Bluetooth connection itself.
+We have to rely on the Bluethoot standard to be secure, as we cannot change it and there are basically no other alternatives to Bluetooth at the moment.
+}
+\item{
+stromschlag bei regen??
+}
+\end{itemize}

### 2015-01-08 proposal/sections/implication-risk.tex
add risks to tex file.

@@
 
-\note{
+\subsection*{implications}
+
+\subsection*{Risks}
+
+\subsubsection*{Less money}
 \begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Importance of the expected results for the discipline
+ \item Description: Get less money to realize the project than defined in the cost accounting.
+ \item Solution: Our project is calculated tightly. So we must revalidate the hole project and create a new proposal with less features and workpackages.
+\end{itemize}
+
+\subsubsection*{More money needed}
 \begin{itemize}
-	\item     To what extent does the proposed research address important challenges?
+ \item Description: The project needs more money than defined in the cost accounting.
+ \item Solution: Looking for new investors to get more money. Sell knowhow to other companies or research groups.
 \end{itemize}
-\item Importance of the expected results for other areas
-\item What are possible risks of the project and how can they be alleviated?
+
+\subsubsection*{Less time for research}
 \begin{itemize}
-	\item     What factors could lead to a failure of the project?
-	\item     Which factors or persons could support the project and increase the chance for success?
-	\item     What if important team members leave the project?
+ \item Description: The brain waves research needs more time than supposed.
+ \item Solution: Buy knowhow from other companies and research groups. Get extern experts to increase the research.
 \end{itemize}
+
+\subsubsection*{Team members leave}
+\begin{itemize}
+ \item Description: One or more team members leave the project.
+ \item Solution: All work topics have two members.
+To avoid loosing knowlegde, there are meetings every week to beware it.
+After knowlegde about leaving a team member, we are looking for a new team member.
 \end{itemize}
-}
+
+\subsubsection*{Brain waves location}
+\begin{itemize}
+ \item  Description: It is not possible to detect brain waves in the near of the ear.
+ \item Solution: We are using sensors to detect brain waves at the forehead.
+\end{itemize}
+
+\subsubsection*{Brain waves detection}
+\begin{itemize}
+ \item Description: It is not possible to detect uniqueness of brain waves.
+ \item Solution: It is not possible to continue, because this is the base of the project.
+\end{itemize}

### 2015-01-08 proposal/sections/team.tex
Costs

@@
 \label{sect:team}
+\subsection{Required Persons and Roles}
+\begin{itemize}
+\item Project Manager
+\item Neuroscience Researcher
+\item Computer Science Researcher
+\item Hardware Engineer
+\item Software Engineer
+\end{itemize}
 
-\note{
+\subsection{Project Manager}
+The project managers have the responsibility of the planning, execution and closing of the project.
+A project manager is the person responsible for accomplishing the stated project objectives.
+Key project management responsibilities include creating clear and attainable project objectives, building the project requirements, and managing the constraints of the project management triangle, which are cost, time, scope and quality.
+He is the bridging gap between the production team and client.
+So he/she must have a fair knowledge of the industry they are in so that they are capable of understanding and discussing the problems with either party.
+
+However, there are some responsibilities that are common to all project Managers, noting:
 \begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Description of the team that is needed to carry out the project. (For the execution phase of the project, not the planning phase.)
-\item How many people?
-\item To what extent are individual members needed?
-\item What knowledge, skills, and experiences are needed for each member?
-\item Demonstrate that the members will be able to carry out the project successfully.
-\item Work structure
+\item Developing the project plan
+\item Managing the project stakeholders
+\item Managing Communication
+\item Managing the project team
+\item Managing the project risk
+\item Managing the project schedule
+\item Managing the project budget
+\item Managing the project conflicts
+\item Managing the project delivery
+\end{itemize}
+So the project manager needs the following skills:
 \begin{itemize}
-	\item     Who will lead the project?
-	\item     How do they work together?
-	\item     Management and coordination
+\item Knowledge of project management
+\item General management knowledge
+\item Product-specific knowledge
+\item Stamina and Endurance
+\item A holistic and sustainable way of thinking
+\item Interpersonal and communication skills
+\end{itemize}
+
+\subsection{Neuroscience Researcher}
+The main tasks of the neuroscience researcher are to prove what kind of brain waves produce recognizable patterns of the same imagination, under what circumstances brain wave scans look similar and to research if the pattern changes if the context of the person changes, like a noisy environment, strong emotions or the effect of drugs.
+The neuroscience researcher works in the first time period of the project, from the 05.01.2015, where the basic research about brain waves should be done in order to achieve the expected results.
+There are two main scientific questions that need to be answered by the neuroscience researcher.
 \begin{itemize}
-		\item 	        What communication structures will be established? (e.g., mailing list, blog, CMS, CVS, ...)
-		\item 	        How often will meetings take place? (Who will participate?)
-		\item 	        How will the work be documented?
-		\item 	        How will information be stored and shared?
+\item{
+First of all he should confirm or disprove the uniqueness of brain waves or define how specific brain waves of a person can be identified.
+}
+\item{
+The second task of the researcher is to confirm the uniqueness of a persons brain waves compared with the brain waves of others.
+Another important factor is to confirm that the brain waves of a specific thought of a person are distinguishable from other thoughts of the same person.
+}
+\item{
+The prototype Wavy should make it possible to indirectly watch the brains function. The activity of the neurons generates an electric field, which can be measured from  outside the skull.
+}
 \end{itemize}
+
+\subsection{Hardware Engineer}                         	
+The hardware engineer begins his work at the beginning of the first phase of the project and starts simultaneously with the neuroscience researcher.
+There already are lots of existing sensors for detecting brain waves.
+The main task of the hardware engineer is to look if a suitable device already exists and can be made smaller or if he has to develop a new device.
+He is responsible for the development of the prototype in order to show that authentication over brain waves works.
+Also, he needs technical knowledge about signal processing.
+
+The goal of his work is to develop a small as possible prototype device, which makes it possible for the user to wear it in the everyday life.
+For that purpose, the project is in need of small as possible sensors, which will be developed by the hardware engineer during his work.
+
+He also needs to develop a micro-controller, which manages the detection and identification of brain waves.
+It is working with the developed algorithm and sending it over a Bluetooth interface to a chosen device.
+The communication with the controller is handled with the developed software interface.
+Therefore, the software engineer and the hardware engineer are working closely together in the same time period.
+
+\subsection{Computer Science Researcher}
+The computer researcher identifies a persons thoughts and reviews existing algorithms for detecting patterns in measured brain waves.
+Since it is possible that the same thought of a person can be shown different in the brain wave, we need some patterns to confirm the measured brain waves implemented in a working algorithm with good performance.
+If there are not found existing algorithms the development of new ones is the goal of the computer science researcher.
+His tasks is also to compare the results with existing authentication methods.
+
+The computer science researcher should have knowledge of machine learning, pattern detection, human computer interface design and signal processing.
+He should have worked in research projects already, where he has experienced some work with brain wave research systems.
+
+\subsection{Software Engineer}
+He is a person concerned with facets of the software development process.
+A software  developer may take part in design, computer programming, or software project management.
+They may contribute to the overview of the project on the application level rather than  component-level or individual programming tasks.
+His task in the project is to work on the communication protocols and the interface of the Wavy device.
+
+This interface should make it possible for the operation system or other developers to handle their authentication over the Wavy.
+He is also responsible for the prototype software, which is used by the WaveMeIn.
+The software engineers working area may include:
+\begin{itemize}
+\item Software design
+\item Implementation
+\item Requirement analysis
+\item {Testing, including defining/supporting acceptance testing and gathering feedback from pre-release testers}
+\item {Participation in software release and post-release activities, including support for product launch evangelism}
+\item {Maintenance}
 \end{itemize}
-\item Cooperations
+
+\subsection{Quality Assurance Manager}
+The quality assurance manager has tasks in the following areas:
 \begin{itemize}
-	\item     Will external cooperators be part of the project? (e.g., other research institutions or companies)
-	\item     What is their role?
-	 \item    Why are they needed?
+\item Controlling
+\item Testing
+\item Reviews
+\item Developing and evaluating statistics
 \end{itemize}
-\end{itemize}
-}
+
+Controlling - the quality assurance manager controls and monitors the progress of the process, he checks if plans are realistic and if a functioning project controlling exists.
+
+Testing - the quality assurance manager checks if the test plan is conform with the quality assurance requirements, he checks if the planed tests have been executed and he monitors the test metrics.
+
+Reviews - the quality assurance manager has to have review-knowledge.
+In reviews he often plays the role of a facilitator, because he checks if the found errors have been corrected after the review process.
+He also checks if the planed reviews have been executed and he is responsible for the quality of the executed review.
+
+\subsection{Work Structure}
+The leader of the project is the project manager, he is responsible for the achievement of project goals, resource goals and timing goals.
+He is also responsible for the management and coordination of the work.
+The team communicates via mailing lists, scrum board and GIT.
+Weekly there will be SCRUM meetings and daily mini scrum meetings with duration of max.
+15 minutes.
+Once a month meetings with all project members will be held and reviews and the work progress will be discussed.The work and the information will be stored and shared with GIT.
+External cooperators will not be part of the project.

### 2015-01-08 proposal/sections/implication-risk.tex
add some text to risks.

@@
 
-\subsection*{implications}
+\subsection*{Implications and challenges}
+One of the biggest challenge in the project is the basic research to detect unique brain waves.
+When our research has success and we are able to detect it, there are many use cases in other topics.
+One use case could be usage for high secure applications and systems.
+Where biometrical methods has some disadvantages, we have discussed above, our method can reduce this disadvantages.
+Other topics could be the medicine. Current we know only a little bit about brain waves. With our research we could increase our medical understanding of brain waves.
 
@@
 \end{itemize}
+
+\subsubsection*{Missinterpretion}
+\begin{itemize}
+ \item Description: The team members not understand the specification correctly.
+ \item Solution: Make meetings and reviews to ensure that every team member has understand it correctly.
+\end{itemize}

### 2015-01-08 proposal/sections/implication-risk.tex
Costs

@@
-\section{Expected implications and risks}
+\section{Expected Implications and Risks}
 \label{sect:implication-risk}
@@
  \item Description: Get less money to realize the project than defined in the cost accounting.
- \item Solution: Our project is calculated tightly. So we must revalidate the hole project and create a new proposal with less features and workpackages.
+ \item Solution: Our project is calculated tightly. So we must revalidate the hole project and create a new proposal with less features and work packages.
 \end{itemize}
@@
  \item Description: The project needs more money than defined in the cost accounting.
- \item Solution: Looking for new investors to get more money. Sell knowhow to other companies or research groups.
+ \item Solution: Looking for new investors to get more money. Sell know-how to other companies or research groups.
 \end{itemize}
@@
  \item Description: The brain waves research needs more time than supposed.
- \item Solution: Buy knowhow from other companies and research groups. Get extern experts to increase the research.
+ \item Solution: Buy know-how from other companies and research groups. Get external experts to increase the research.
 \end{itemize}
@@
  \item Solution: All work topics have two members.
-To avoid loosing knowlegde, there are meetings every week to beware it.
-After knowlegde about leaving a team member, we are looking for a new team member.
+To avoid loosing knowledge, there are meetings every week to beware it.
+After knowledge about leaving a team member, we are looking for a new team member.
 \end{itemize}
@@
 
-\subsubsection*{Missinterpretion}
+\subsubsection*{Misinterpretation}
 \begin{itemize}

### 2015-01-08 diary/compile_diary.sh
diary submission 4

@@ -1,4 +1,4 @@
 #find diary/ -name \*.md -type f -exec pandoc -o {}.tex {} \;
 pandoc -f markdown_phpextra -o diary_bundle.tex reports/*.md
-pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_3.tex
+pdflatex -synctex=1 -interaction=nonstopmode 10_Diary_4.tex
 rm *.out *.log *.synctex.gz *.aux

### 2015-01-08 proposal/sections/costs.tex
Costs
cost updates

@@
 \label{sect:costs}
-\subsection{Person}
-\begin{tabular}{|c|c|c|c|c|c|c|}
+\subsection{Personnel}
+
+\begin{tabular}{|l|c|c|c|c|c|c|}
 \hline 
-Person & Commitment & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
+Person & \% & 1st Year & 2nd Year & 3rd Year & 4th Year & Sum \\ 
 \hline 
-Project Manager & 100\% & 54000 & • & • & • & • \\ 
+Project Manager & 100\% & 54000,00 & 55080,00 & 56181,60 & 28652,62 & 193914,22 \\ 
 \hline 
-Quality Manager & 75\% & 27000 & • & • & • & • \\ 
+Quality Manager & 75\% & 27000,00 & 27540,00 & 28090,80 & 14326,21 & 96957,11 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Neuroscience Researcher 1 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Neuroscience Researcher 2 & 100\% & 41810,00 & 42646,20 & - & - & 84856,20 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Computer Science R. 1 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Computer Science R. 2 & 100\% & 41810,00 & 42646,20 & 43499,12 & 22184,55 & 150139,88 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Hardware Engineer 1 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Hardware Engineer 2 & 100\% & 31160,00 & 31783,20 & 32418,86 & 16533,62 & 111895,67 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Software Engineer 1 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
 \hline 
-• & • & • & • & • & • & • \\ 
+Software Engineer 2 & 100\% & 31160,00 & 31783,20 & - & - & 62943,20 \\ 
+\hline 
+\textbf{SUM} & & 372880,00 & 380337,60 & 236108,38 & 120415,27 & 989325,98 \\ 
 \hline 
@@
 
-\subsection{Hardware}
-\begin{tabular}{|c|c|c|c|c|c|}
+\paragraph{Project Manager}
+A project manager supervises the project from the beginning to the end.
+His role has scheduling and supportive character, works close to everyone an thus has a good grasp of all parts of the project.
+Despite having a good overview, he is not deeply involved in any of the other employers work and not accumulating too specific knowledge.
+This way the risks involving a switch of the project managers position should theoretically be less of an impact then changing an expert.
+
+\paragraph{Quality Manager}
+The quality manager has a supportive work.
+Inspecting, improving and creating new standards during the whole project lifecycle helps raising work and product quality nd avoiding deadline misses.
+As the team is not very big, we think of a 75\% position %< sotz ausformulieren
+
+\paragraph{Neuroscience Researcher}
+The neuroscience researcher has a highly specialized research and develop task at the beginning of the project. To avoid risks and speed up the initial research two neuroscientists get hired.
+
+\paragraph{Computer Science Researcher}
+The computer science researchers care about algorithms and machine learning.
+They will be occupied  during the whole project.
+Two researcher should avoid risks of loosing knowledge if one leaves the project and can speed up the research especially within the first ans second milestone.
+Their field of work overlaps with all other project members.
+
+\paragraph{Hardware Engineer}
+The hardware engineer starts with analyzing existing devices and sensors.
+Working together with neuroscientists they create the Wavy prototype and are therefore part of the project during the whole 42 month.
+Two engineers provide a minimal backup.
+
+\paragraph{Software Engineer}
+The software engineers start implementing the knowledge gathered by neuroscience and computer science researcher starting with the second milestone.
+As for most positions two members seems fit to provide backup.
+
+\subsection{Equipment}
+\begin{tabular}{|l|c|c|c|c|c|}
 \hline 
@@
 \hline 
-Commercial EEG Headsets & 3000 & • & • & • & • \\ 
+Commercial EEG Headsets & 3000,00 & - & - & - & 3000,00 \\ 
 \hline 
-Medical EEG Device & 60000 & • & • & • & • \\ 
+Medical EEG Device & 60000,00 & - & - & - & 60000,00 \\ 
 \hline 
-Server & 4000 & • & • & • & • \\ 
+Server & 4000 & - & - & - & 4000,00 \\ 
 \hline 
-Workstations & Notebooks & 6400 & • & • & • & • \\ 
+Workstations/Notebooks & 6400 & - & - & - & 6400,00 \\ 
+\hline 
+\textbf{SUM} & 77000,00 &  &  &  & 77000,00 \\ 
 \hline 
 \end{tabular}
+
+The equipment, which is needed for the WaveMeInProject will be purchased in the 1st year, since the equipment is required already in the reaserch part of the project and includes following hardware parts:
+hardware Commercial EEG Headsets with costs of overall 3000\euro{} - the price per item is 499\euro{} and the project requires 6 pieces - https://emotiv.com/store/epoc-detail/
+
 \subsection{Material}
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|}
 \hline 
@@
 \hline 
-Hardware for the Prototype & 2000 & • & • & • & 2000 \\ 
+Hardware for the Prototype & 2000 & - & - & - & 2000 \\ 
 \hline 
-EEG Sensors & 500 & • & • & • & 500 \\ 
+EEG Sensors & 500 & - & - & - & 500 \\ 
 \hline 
-• & • & • & • & • & • \\ 
+\textbf{SUM} & 2500 &  &  &  & 2500 \\ 
 \hline 
 \end{tabular}
+
 \subsection{Travel Costs}
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|}
 \hline 
@@
 \hline 
-Conferences & 4000 & • & 4000 & • & 8000 \\ 
+Conferences & 4000 & - & 4000 & - & 8000 \\ 
 \hline 
-EEG Sensors & 500 & • & • & • & 500 \\ 
-\hline 
-• & • & • & • & • & • \\ 
+\textbf{SUM} & 4000 &  & 4000 &  & 8000 \\ 
 \hline 
 \end{tabular}
+
+The travel costs are estimated to be 8000\euro{}, whereof for the 1st and for the 3rd year 4000\euro{} are provided for advanced training courses.
+
 \subsection{Other Costs}
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|}
 \hline 
@@
 \hline 
-Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 8000 \\
+Teambuilding, Snacks, Coffee, Meetings & 7000 & 7000 & 7000 & 3500 & 24500 \\
 \hline 
 \end{tabular}
+
+The other costs for teambuilding, snacks, coffee or meetings are estimated to be in sum 24500\euro{}: 7000\euro{} in the 1st year, 7000\euro{} in the 2nd year, 7000\euro{} in the 3rd year and 3500\euro{} in the 4th year.

### 2015-01-08 proposal/sections/ethics-security.tex

--- 2014-11-24 proposal/sections/ethics-security.tex
@@
-\section{Ethical considerations \& security issues}
+\section{Ethical Considerations \& Security Issues}
 \label{sect:ethics-security}
+\subsection{Ethical Considerations}
 
-\note{
-\begin{itemize}
-\item {\em Length: 1-2 pages}
-\item Provide a brief explanation of the ethical issue involved and how it will be dealt with appropriately.
-\item Are there any security-sensitive issues that apply to your proposal?
-\end{itemize}
-}
+\subsubsection*{Electromagnetic Waves}
+At the moment there is an ongoing dispute in the public over weather or not long-term exposure to electromagnetic fields can cause cancer or changes in the brain chemistry.
+As far as we know today weak electromagnetic fields will have no effects on the human brain, but strong fields definitively do.
+The Wavy device will definitely create such a weak magnetic field (like all electric devices) and due to its location at the users ear and the intention of wearing it permanently it has the potential of harming the user if electromagnetic waves really have negative effects on the human brain.
+
+\subsubsection*{Collecting User Data}
+The Wavy device could be copied by companies trying to make a profit by constantly monitoring the emotional state of the user.
+The device would still function as an access device, but it could be misused for marketing purposes for example.
+A possible scenario could be: A company installs a wireless module into a TV set that sends a signal when an advertisement is shown.
+This signal activates an application in the user’s phone which in turn starts collecting emotional data via such a “bad” Wavy device from the user who is watching the advertisement.
+The company can then sell the gathered knowledge about the user’s response to the advertisement and his/her preferences to others such as the advertising company.
+
+\subsection{Security Issues}
+\subsubsection*{Mind Reading}
+At the current state of the technology mind reading is still science fiction, but basically a device such as the Wavy could be used to read all thoughts of a user.
+The only real issue that is preventing this scenario at the moment is that from a given set of brain waves there is currently no way of translating this user’s thought into a visual or textual representation for someone else to read.
+However, this may be possible at some point in the future.
+A compromised Wavy device poses a serious security risk for the user.
+
+\subsubsection*{Bluetooth}
+The biggest issue in security is probably the Bluetooth connection itself.
+We have to rely on the Bluetooth standard to be secure, as we cannot change it and there are basically no other alternatives to Bluetooth at the moment.
